---
title: "Acoustic Analysis"
author: "Tabea Thies"
date: "`r Sys.Date()`"
output: html_document
---

# Comparing levodopa and DBS effect on acoustic vowel durations

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load necessary packages

```{r load_tidyverse, warning = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(ggbeeswarm)
library(ggsignif)
```

Load data set

```{r load_counts, message = FALSE, warning = FALSE}
data <- read_csv('../data/acoustics.csv')
```

Create subset for the comparison med-OFF vs med-ON vs DBS-ON

med-OFF functions as baseline here

```{r, message = FALSE, warning = FALSE}
data_red <-  filter(data, med_con %in% c('med-OFF', 'med-ON', 'med-OFF/DBS-ON'))

```

Descriptive stats for reporting values in Section 9.2.2

```{r, message = FALSE, warning = FALSE}

V1_stats <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(V1.dur)),
            SD = round(sd(V1.dur)))
V1_stats 

V1_stats2 <- data_red %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(V1.dur)),
            SD = round(sd(V1.dur)))
V1_stats2 

```

Plot means per group and focus condition (Figure 9.1)

```{r, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 6}

DUR_plot <- data_red %>% 
  group_by(med_con, speaker) %>% 
  summarise(M = round(mean(V1.dur)),
            SD = round(sd(V1.dur)))

plot01 <- DUR_plot %>%
  ggplot(aes(x=factor(med_con, level = c("med-ON", "med-OFF", "med-OFF/DBS-ON")), y=M, fill=med_con)) +
  geom_boxplot(width=0.4, color="black", alpha=0.1, outlier.alpha = 0.1) +
  stat_boxplot(geom = "errorbar", width = 0.1) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=1)+
  theme_classic() +
  theme(legend.position = 'none')+
  scale_fill_manual(values = c('#BE7592', "#AD8A52", '#74BDA0')) +
  xlab(NULL) +
  scale_x_discrete(labels = c('med-ON','baseline', "DBS-ON"))+
  ylab("Vowel Duration (ms)")+
  coord_cartesian(ylim = c(70, 225)) + 
  scale_y_continuous(breaks = seq(80, 200, 40)) +
  theme(axis.title.y = element_text(face = 'bold', size = 30, margin = margin(r = 30)))+
  theme(strip.text.x = element_text(size=32, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 28),
        axis.text.y = element_text(size = 28)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())+
  geom_signif(comparisons=list(c("med-OFF", "med-OFF/DBS-ON")), annotations="p < .001", y_position = 214)+
  geom_signif(comparisons=list(c("med-ON", "med-OFF/DBS-ON")), annotations="p < .001", y_position = 220)

ggsave(plot01, filename = '../plots/V1dur_DBS2.png',
       width = 10, height = 7, dpi = 300)

```

# Statistical analysis

Load packages needed

```{r, message = FALSE, warning = FALSE}
library(lme4)
library(lmerTest)
library(sjPlot)
library(sjmisc)
library(merTools)
library(emmeans)

set.seed(42)
```

Create model with lmer to test change in vowel duration form baseline (med-OFF) to DBS-ON and med-ON

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(V1.dur ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

mdl2 <- lmer(V1.dur ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(V1.dur ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2b) ## effect of focus: X2(2) = 26.756, p = 1.549e-06

emmeans(mdl2, specs = pairwise ~ focus_con)

## check group

mdl2c <- lmer(V1.dur ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2c) ##  group effect: X2(1) = 49.16, p = 2.114e-11 

emmeans(mdl2, specs = pairwise ~ med_con)
```