---
title: "Acoustic Analysis"
author: "Tabea Thies"
date: "`r Sys.Date()`"
output: html_document
---

# Comparing levoDOPA and DBS effect on acoustic vowel durations

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load necessary packages

```{r load_tidyverse, warning = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
```

Load data set

```{r load_counts, message = FALSE, warning = FALSE}
data <- read_csv('../data/acoustics.csv')
```

Create subset for the comparison med-OFF vs med-ON vs DBS-ON

med-OFF functions as baseline here

```{r, message = FALSE, warning = FALSE}
data_red <-  filter(data, med_con %in% c('med-OFF', 'med-ON', 'med-OFF/DBS-ON'))

```

Descriptive stats for reporting values in Section 9.2.2

```{r, message = FALSE, warning = FALSE}

V1_stats <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(V1.dur)),
            SD = round(sd(V1.dur)))
V1_stats 

V1_stats2 <- data_red %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(V1.dur)),
            SD = round(sd(V1.dur)))
V1_stats2 

```

Plot means per group and focus condition (Figure 9.1)

```{r, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 6}

V1_stats2 %>% 
  ggplot(aes(x=factor(med_con, level = c("med-ON", "med-OFF", "med-OFF/DBS-ON")), y=M, group = focus_con, col=focus_con)) + 
  geom_line(size = 1.5) +
  geom_point(size = 10)+
  theme_classic() +
  theme(legend.position = 'right')+
  theme(legend.title = element_blank(),
        legend.text = element_text(size = 28))+
  scale_color_manual(values = c("#8da0cb", "#66c2a5", "#fc8d62")) +
  xlab(NULL) +
  scale_x_discrete(labels = c('med-ON','baseline', "DBS-ON"))+
  ylab("Vowel Duration (ms)")+
  coord_cartesian(ylim = c(115, 150)) + 
  scale_y_continuous(breaks = seq(120, 150, 10)) +
  theme(axis.title.y = element_text(face = 'bold', size = 30, margin = margin(r = 30)))+
  theme(strip.text.x = element_text(size=32, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 28),
        axis.text.y = element_text(size = 28)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())

```

# Statistical analysis

Load packages needed

```{r, message = FALSE, warning = FALSE}
library(lme4)
library(lmerTest)
library(sjPlot)
library(sjmisc)
library(merTools)
```

Create model with lmer to test change in vowel duration form baseline (med-OFF) to DBS-ON

```{r, message = FALSE, warning = FALSE}

data_red2 <-  filter(data, med_con %in% c('med-OFF', 'med-OFF/DBS-ON'))

set.seed(42)

mdl <- lmer(V1.dur ~ med_con + focus_con + vowel + cons + age + (1 + focus_con + med_con|speaker), data_red2, REML = FALSE)
```

Look at model summary

```{r, message = FALSE, warning = FALSE}
summary(mdl)
```

Extracting fit statistics for reporting in text

```{r, message = FALSE, warning = FALSE}
tab_model(mdl)
```

## Explore random effects with function REsim
Effects for individual levels are simulated\
Output: mean, median and sd of the random effect estimates\

Plots estimated random effects for each participant\
95 % Intervals that do not overlap zero are bold

```{r, message = FALSE, warning = FALSE}
randoms <- REsim(mdl) 
plotREsim(randoms)

```

Plot estimated 95 % interval spans for the intercept of acoustic vowel duration

```{r, message = FALSE, warning = FALSE}

random2 <- subset(randoms, term == "(Intercept)")

random2 %>% 
  ggplot(aes(x = mean, xmin = mean-sd-sd, xmax = mean+sd+sd, y = reorder(groupID, mean))) +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  geom_point(size = 5, color = '#AD8A52')+
  geom_errorbar(size = 1) +
  theme_classic()+
  xlab('95 % interval spans of estimated random intercepts') +
  ylab('Speaker')+
  theme(axis.title.y = element_text(face = 'bold', size = 18, margin = margin(r = 12)))+
  theme(axis.title.x = element_text(face = 'bold', size = 18, margin = margin(t = 12)))+
  theme(strip.text.x = element_text(size=18, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())+
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 18))
```

Plot estimated 95 % interval spans for the slopes of acoustic vowel duration

```{r, message = FALSE, warning = FALSE}

random3 <- subset(randoms, term == "med_conmed-OFF/DBS-ON")

random3 %>% 
  ggplot(aes(x = mean, xmin = mean-sd-sd, xmax = mean+sd+sd, y = reorder(groupID, mean))) +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  geom_point(size = 5, color = '#AD8A52')+
  geom_errorbar(size = 1) +
  theme_classic()+
  xlab('95 % interval spans of estimated random slopes') +
  ylab('Speaker')+
  theme(axis.title.y = element_text(face = 'bold', size = 18, margin = margin(r = 12)))+
  theme(axis.title.x = element_text(face = 'bold', size = 18, margin = margin(t = 12)))+
  theme(strip.text.x = element_text(size=18, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())+
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 18))
```



