---
title: "Articulatory Analysis"
author: "Tabea Thies"
date: "`r Sys.Date()`"
output: html_document
---

# Comparing levodopa and DBS effect on tongue body movements

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load necessary packages

```{r load_tidyverse, warning = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(merTools)
library(sjPlot)
library(sjmisc)
library(emmeans)
library(ggsignif)

set.seed(42)
```

Load data set

```{r load_counts, message = FALSE, warning = FALSE}
data <- read_csv('../data/articulation.csv')
```

Create subset for the comparison med-OFF vs med-ON vs DBS-ON

med-OFF functions as baseline here

```{r, message = FALSE, warning = FALSE}

data_red <-  filter(data, med_con %in% c('med-OFF', 'med-ON', 'med-OFF/DBS-ON'))

```

# Movement duration

Descriptive stats for reporting values in Section 9.2.3

```{r, message = FALSE, warning = FALSE}

DUR_stats <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(GAI_V1)),
            SD = round(sd(GAI_V1)))
DUR_stats 

DUR_stats2 <- data_red %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(GAI_V1)),
            SD = round(sd(GAI_V1)))
DUR_stats2

```

Plot tongue body movement duration

```{r, message = FALSE, warning = FALSE}

DUR_plot <- data_red %>% 
  group_by(med_con, speaker) %>% 
  summarise(M = round(mean(GAI_V1)),
            SD = round(sd(GAI_V1)))

plot01 <- DUR_plot %>%
  ggplot(aes(x=factor(med_con, level = c("med-ON", "med-OFF", "med-OFF/DBS-ON")), y=M, fill=med_con)) +
  geom_boxplot(width=0.4, color="black", alpha=0.1, outlier.alpha = 0.1) +
  stat_boxplot(geom = "errorbar", width = 0.1) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=1)+
  theme_classic() +
  theme(legend.position = 'bottom')+
  theme(legend.title = element_blank(),
        legend.text = element_text(size = 38))+
  scale_fill_manual(labels = c("med-ON", "baseline", "DBS-ON"), 
                    breaks = c("med-ON", "med-OFF", "med-OFF/DBS-ON"),
                    values = c('#74BDA0', '#BE7592', "#AD8A52")) +
  xlab(NULL) +
  ylab("Duration (ms)")+
  coord_cartesian(ylim = c(135, 312)) + 
  #scale_y_continuous(breaks = seq(80, 220, 40)) +
  theme(axis.title.y = element_text(face = 'bold', size = 30, margin = margin(r = 30)))+
  theme(strip.text.x = element_text(size=32, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 28)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())+
  geom_signif(comparisons=list(c("med-OFF", "med-ON")), annotations="p = .009", y_position = 295)+
  geom_signif(comparisons=list(c("med-ON", "med-OFF/DBS-ON")), annotations="p < .001", y_position = 307)

ggsave(plot01, filename = '../plots/TB_DUR_DBS2.png',
       width = 10, height = 7, dpi = 300)
        
```

Create linear mixed model to investigate effect on TB duration

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(GAI_V1 ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

mdl2 <- lmer(GAI_V1 ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus
mdl2b <- lmer(GAI_V1 ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2b) ## effect of focus: X(2) = 13.897, p = 0.0009603

emmeans(mdl2, specs = pairwise ~ focus_con)

## check group

mdl2c <- lmer(GAI_V1 ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2c) ## group effect: X(2) = 23.495, p = 7.911e-06

emmeans(mdl2, specs = pairwise ~ med_con)

```

# Movement amplitude

Descriptive stats for reporting values in Section 10.2.3

```{r, message = FALSE, warning = FALSE}

DISPL_stats <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(disp_V1_abs),1),
            SD = round(sd(disp_V1_abs),1))
DISPL_stats 

DISPL_stats2 <- data_red %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(disp_V1_abs),1),
            SD = round(sd(disp_V1_abs),1))
DISPL_stats2

```

Plot tongue body movement amplitude

```{r, message = FALSE, warning = FALSE}

DISPL_plot <- data_red %>% 
  group_by(med_con, speaker) %>% 
  summarise(M = round(mean(disp_V1_abs),1),
            SD = round(sd(disp_V1_abs),1))

plot02 <- DISPL_plot %>%
  ggplot(aes(x=factor(med_con, level = c("med-ON", "med-OFF", "med-OFF/DBS-ON")), y=M, fill=med_con)) +
  geom_boxplot(width=0.4, color="black", alpha=0.1, outlier.alpha = 0.1) +
  stat_boxplot(geom = "errorbar", width = 0.1) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=1)+
  theme_classic() +
  theme(legend.position = 'bottom')+
  theme(legend.title = element_blank(),
        legend.text = element_text(size = 38))+
  scale_fill_manual(labels = c("med-ON", "baseline", "DBS-ON"), 
                    breaks = c("med-ON", "med-OFF", "med-OFF/DBS-ON"),
                    values = c('#74BDA0', '#BE7592', "#AD8A52")) +
  xlab(NULL) +
  ylab("Amplitude (mm)")+
  coord_cartesian(ylim = c(3.5, 11.5)) + 
  scale_y_continuous(breaks = seq(6, 10, 2)) +
  theme(axis.title.y = element_text(face = 'bold', size = 30, margin = margin(r = 30)))+
  theme(strip.text.x = element_text(size=32, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 28)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())+
  geom_signif(comparisons=list(c("med-OFF", "med-OFF/DBS-ON")), annotations="p < .001", y_position = 10.9)+
  geom_signif(comparisons=list(c("med-ON", "med-OFF/DBS-ON")), annotations="p = .017", y_position = 11.3)


ggsave(plot02, filename = '../plots/TB_DISPL_DBS2.png',
       width = 10, height = 7, dpi = 300)
        
```

Create linear mixed model to investigate effect on TB amplitude

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(disp_V1_abs ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

mdl2 <- lmer(disp_V1_abs ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus
mdl2b <- lmer(disp_V1_abs ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2b) ## effect of focus: x(2) = 7.3794, p = 0.02498

emmeans(mdl2, specs = pairwise ~ focus_con)

## check group

mdl2c <- lmer(disp_V1_abs ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2c) ## group effect: X(1) = 18.421, p = 9.998e-05

emmeans(mdl2, specs = pairwise ~ med_con)

```

# Movement peak velocities

Descriptive stats for reporting values in Section 10.2.3

```{r, message = FALSE, warning = FALSE}

PVEL_stats <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(pvel_V1_abs)),
            SD = round(sd(pvel_V1_abs)))
PVEL_stats 

PVEL_stats2 <- data_red %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(pvel_V1_abs)),
            SD = round(sd(pvel_V1_abs)))
PVEL_stats2

```

Plot tongue body movement velocity

```{r, message = FALSE, warning = FALSE}

PVEL_plot <- data_red %>% 
  group_by(med_con, speaker) %>% 
  summarise(M = round(mean(pvel_V1_abs),2),
            SD = round(sd(pvel_V1_abs),2))

plot03 <- PVEL_plot %>%
  ggplot(aes(x=factor(med_con, level = c("med-ON", "med-OFF", "med-OFF/DBS-ON")), y=M, fill=med_con)) +
  geom_boxplot(width=0.4, color="black", alpha=0.1, outlier.alpha = 0.1) +
  stat_boxplot(geom = "errorbar", width = 0.1) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=1)+
  theme_classic() +
  theme(legend.position = 'bottom')+
  theme(legend.title = element_blank(),
        legend.text = element_text(size = 38))+
  scale_fill_manual(labels = c("med-ON", "baseline", "DBS-ON"), 
                    breaks = c("med-ON", "med-OFF", "med-OFF/DBS-ON"),
                    values = c('#74BDA0', '#BE7592', "#AD8A52")) +
  xlab(NULL) +
  ylab("Peak Velocity (mm/s)")+
  coord_cartesian(ylim = c(30, 115)) + 
  scale_y_continuous(breaks = seq(40, 100, 20)) +
  theme(axis.title.y = element_text(face = 'bold', size = 30, margin = margin(r = 30)))+
  theme(strip.text.x = element_text(size=32, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 28)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())+
  geom_signif(comparisons=list(c("med-OFF", "med-OFF/DBS-ON")), annotations="p < .001", y_position = 109)+
  geom_signif(comparisons=list(c("med-ON", "med-OFF/DBS-ON")), annotations="p < .001", y_position = 113)


ggsave(plot03, filename = '../plots/TB_PVEL_DBS2.png',
       width = 10, height = 7, dpi = 300)
        
```

Create linear mixed model to investigate disease effect on TB velocity

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(pvel_V1_abs ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

mdl2 <- lmer(pvel_V1_abs ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(pvel_V1_abs ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2b) ## no effect of focus: X(2) = 0.9796, p = 0.6128

## check group

mdl2c <- lmer(pvel_V1_abs ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2c) ## group effect: X(1) = 38.687, p = 3.974e-09 

emmeans(mdl2, specs = pairwise ~ med_con)

```

# Create Multiplot for duration & amplitude

```{r, message = FALSE, warning = FALSE}

library(grid)
library(gridExtra)
library(cowplot)
library(lemon)

ggsave(filename = "../plots/TB_DUR_DISPL_PVEL_DBS2.png", 
       grid_arrange_shared_legend(plot01, plot02, plot03, ncol=3, nrow=1,
                    top = textGrob(" \nTongue Body Movement\n ",gp=gpar(fontsize=40,font="bold", margin(b = 20)))),
       width = 25, height = 10, dpi = 300, units = "in", device='png')

```

# Movement Profiles

Descriptive stats for reporting values in Table 7.2

Subset for production of vowels /a/ and /i/

```{r, message = FALSE, warning = FALSE}

data_red_i_a <-  filter(data_red, vowel %in% c('i', 'a'))

```

Acceleration phase

```{r, message = FALSE, warning = FALSE}

ACC_stats <- data_red_i_a %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(acc_V1)),
            SD = round(sd(acc_V1)))
ACC_stats 

ACC_stats2 <- data_red_i_a %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(acc_V1)),
            SD = round(sd(acc_V1)))
ACC_stats2

```

Deceleration phase

```{r, message = FALSE, warning = FALSE}

DEC_stats <- data_red_i_a %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(dec_V1)),
            SD = round(sd(dec_V1)))
DEC_stats 

DEC_stats2 <- data_red_i_a %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(dec_V1)),
            SD = round(sd(dec_V1)))
DEC_stats2

```

Symmetry profile

```{r, message = FALSE, warning = FALSE}

SYMM_stats <- data_red_i_a %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(symmetry_V1),1),
            SD = round(sd(symmetry_V1),1))
SYMM_stats 

SYMM_stats2 <- data_red_i_a %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(symmetry_V1),1),
            SD = round(sd(symmetry_V1),1))
SYMM_stats2

```

Create linear mixed model to investigate disease effect on symmetry

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(symmetry_V1 ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red_i_a, REML = FALSE)

mdl2 <- lmer(symmetry_V1 ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red_i_a, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(symmetry_V1 ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red_i_a, REML = FALSE)

anova(mdl2, mdl2b) ## no effect of focus: X(2) = 2.1241, p = 0.3457

## check group

mdl2c <- lmer(symmetry_V1 ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red_i_a, REML = FALSE)

anova(mdl2, mdl2c) ## no group effect: X(1) = 0.5536, p = 0.7582

```

Stiffness

```{r, message = FALSE, warning = FALSE}

STIFF_stats <- data_red_i_a %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(stiffness_V1_abs),1),
            SD = round(sd(stiffness_V1_abs),1))
STIFF_stats 

STIFF_stats2 <- data_red_i_a %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(stiffness_V1_abs),1),
            SD = round(sd(stiffness_V1_abs),1))
STIFF_stats2

```

Create linear mixed model to investigate disease effect on stiffness

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(stiffness_V1_abs ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red_i_a, REML = FALSE)

mdl2 <- lmer(stiffness_V1_abs ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red_i_a, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(stiffness_V1_abs ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red_i_a, REML = FALSE)

anova(mdl2, mdl2b) ## no effect of focus: X(2) = 3.0226, p = 0.2206

## check group

mdl2c <- lmer(stiffness_V1_abs ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red_i_a, REML = FALSE)

anova(mdl2, mdl2c) ## no group effect: X(1) = 4.39 , p = 0.1114

```

# Timing pattern

```{r, message = FALSE, warning = FALSE}

ONS_avgs <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(onsC1_to_onsV1)),
            SD = round(sd(onsC1_to_onsV1)))
ONS_avgs

TARG_avgs <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(targC1_to_targV1)),
            SD = round(sd(targC1_to_targV1)))
TARG_avgs

OnsAlign_avgs <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(onsV1.syll.dur),2)*100,
            SD = round(sd(onsV1.syll.dur),2)*100)
OnsAlign_avgs

TargAlign_avgs <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(targV1.syll.dur),2)*100,
            SD = round(sd(targV1.syll.dur),2)*100)
TargAlign_avgs

```

Statistical analysis for timing patterns

create model for onset to onset interval

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(onsC1_to_onsV1 ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

mdl2 <- lmer(onsC1_to_onsV1 ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(onsC1_to_onsV1 ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2b) ## no effect of focus: X(2) = 2.2727, p = 0.321

## check group

mdl2c <- lmer(onsC1_to_onsV1 ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2c) ## no group effect: X(1) = 3.3959, p = 0.1831

```

create model for target to target interval

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(targC1_to_targV1 ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

mdl2 <- lmer(targC1_to_targV1 ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(targC1_to_targV1 ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2b) ## effect of focus: X(2) = 17.97, p = 0.0001253

emmeans(mdl2, specs = pairwise ~ focus_con)

## check group

mdl2c <- lmer(targC1_to_targV1 ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2c) ## group effect: X(2) = 15.573, p = 0.0004153

emmeans(mdl2, specs = pairwise ~ med_con)

```

create model for onset alignment

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(onsV1.syll.dur ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

mdl2 <- lmer(onsV1.syll.dur ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(onsV1.syll.dur ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2b) ## effect of focus: X(2) = 10.917 , p = 0.004259

emmeans(mdl2, specs = pairwise ~ focus_con)

## check group

mdl2c <- lmer(onsV1.syll.dur ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2c) ## group effect: X(1) = 6.6033, p = 0.03682

emmeans(mdl2, specs = pairwise ~ med_con)

```

create model for target alignment

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(targV1.syll.dur ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

mdl2 <- lmer(targV1.syll.dur ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus
mdl2b <- lmer(targV1.syll.dur ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2b) ## no effect of focus: X(2) = 0.1418, p = 0.9315

## check group

mdl2c <- lmer(targV1.syll.dur ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2c) ## no group effect: X(1) = 1.1284, p = 0.5688
```