---
title: "Articulatory Analysis"
author: "Tabea Thies"
date: "`r Sys.Date()`"
output: html_document
---

# Analysis of tongue body movements comparing healthy controls and speakers with PD (med-OFF)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load necessary packages

```{r load_tidyverse, warning = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(merTools)
library(sjPlot)
library(sjmisc)
library(emmeans)

set.seed(42)
```

Load data set

```{r load_counts, message = FALSE, warning = FALSE}
data <- read_csv('../data/articulation.csv')
```

Create subset for the comparison Controls vs PD (in med-OFF)

```{r, message = FALSE, warning = FALSE}
data_red <-  filter(data, med_con %in% c('Control', 'med-OFF'))
```

# Movement duration

Descriptive stats for reporting values in Section 6.2.3

```{r, message = FALSE, warning = FALSE}

DUR_stats <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(GAI_V1)),
            SD = round(sd(GAI_V1)))
DUR_stats 

DUR_stats2 <- data_red %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(GAI_V1)),
            SD = round(sd(GAI_V1)))
DUR_stats2

```

Plot tongue body movement duration

```{r, message = FALSE, warning = FALSE}

group_names <- as_labeller(
     c(`Control` = "CON", `med-OFF` = "PD"))

DUR_plot <- data_red %>% 
  group_by(med_con, focus_con, speaker) %>% 
  summarise(M = round(mean(GAI_V1)),
            SD = round(sd(GAI_V1)))

plot01 <- DUR_plot %>%
  ggplot(aes(x=focus_con, y=M, fill=focus_con)) +
  geom_boxplot(width=0.4, color="black", alpha=0.1, outlier.alpha = 0.1) +
  stat_boxplot(geom = "errorbar", width = 0.1) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=1)+
  theme_classic() +
  theme(legend.position = 'bottom')+
  theme(legend.title = element_blank(),
        legend.text = element_text(size = 38))+
  scale_fill_manual(values = c("#8da0cb", "#66c2a5", "#fc8d62")) +
  xlab(NULL) +
  facet_grid(.~med_con, labeller = group_names)+
  ylab("Duration (ms)")+
  coord_cartesian(ylim = c(135, 310)) + 
  scale_y_continuous(breaks = seq(150, 300, 50)) +
  theme(axis.title.y = element_text(face = 'bold', size = 34, margin = margin(r = 30)))+
  theme(strip.text.x = element_text(size=32, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 28)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())

ggsave(plot01, filename = '../plots/TB_DUR_CON_PD.png',
       width = 12, height = 7, dpi = 300)
        
```

Create linear mixed model to investigate disease effect on TB duration

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(GAI_V1 ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

mdl2 <- lmer(GAI_V1 ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(GAI_V1 ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2b) ## effect of focus: X2(2) = 25.423, p = 3.017e-06

emmeans(mdl2, specs = pairwise ~ focus_con)

## check group

mdl2c <- lmer(GAI_V1 ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2c) ## group effect: X2(1) = 5.8087, p = 0.01595 *

emmeans(mdl2, specs = pairwise ~ med_con)

```

# Movement amplitude

Descriptive stats for reporting values in Section 6.2.3

```{r, message = FALSE, warning = FALSE}

DISPL_stats <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(disp_V1_abs),1),
            SD = round(sd(disp_V1_abs),1))
DISPL_stats 

DISPL_stats2 <- data_red %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(disp_V1_abs),1),
            SD = round(sd(disp_V1_abs),1))
DISPL_stats2

```

Plot tongue body movement amplitude

```{r, message = FALSE, warning = FALSE}

DISPL_plot <- data_red %>% 
  group_by(med_con, focus_con, speaker) %>% 
  summarise(M = round(mean(disp_V1_abs),1),
            SD = round(sd(disp_V1_abs),1))

plot02 <- DISPL_plot %>%
  ggplot(aes(x=focus_con, y=M, fill=focus_con)) +
  geom_boxplot(width=0.4, color="black", alpha=0.1, outlier.alpha = 0.1) +
  stat_boxplot(geom = "errorbar", width = 0.1) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=1)+
  theme_classic() +
  theme(legend.position = 'bottom')+
  theme(legend.title = element_blank(),
        legend.text = element_text(size = 38))+
  scale_fill_manual(values = c("#8da0cb", "#66c2a5", "#fc8d62")) +
  xlab(NULL) +
  facet_grid(.~med_con, labeller = group_names)+
  ylab("Amplitude (mm)")+
  coord_cartesian(ylim = c(3, 12.5)) + 
  scale_y_continuous(breaks = seq(4, 12, 4)) +
  theme(axis.title.y = element_text(face = 'bold', size = 34, margin = margin(r = 30)))+
  theme(strip.text.x = element_text(size=32, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 28)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())

ggsave(plot02, filename = '../plots/TB_DISPL_CON_PD.png',
       width = 12, height = 7, dpi = 300)

``` 

Create linear mixed model to investigate disease effect on TB amplitude

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(disp_V1_abs ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

mdl2 <- lmer(disp_V1_abs ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(disp_V1_abs ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2b) ## effect of focus: X2(2) = 10.968, p = 0.004153

emmeans(mdl2, specs = pairwise ~ focus_con)

## check group

mdl2c <- lmer(disp_V1_abs ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2c) ## no group effect: X2(1) = 0.6644, p = 0.415

```

# Movement peak velocities

Descriptive stats for reporting values in Section 6.2.3

```{r, message = FALSE, warning = FALSE}

PVEL_stats <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(pvel_V1_abs)),
            SD = round(sd(pvel_V1_abs)))
PVEL_stats 

PVEL_stats2 <- data_red %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(pvel_V1_abs)),
            SD = round(sd(pvel_V1_abs)))
PVEL_stats2

```

Create linear mixed model to investigate disease effect on TB velocity

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(pvel_V1_abs ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

mdl2 <- lmer(pvel_V1_abs ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(pvel_V1_abs ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2b) ## no effect of focus: X(2) = 1.9891, p = 0.3699

## check group

mdl2c <- lmer(pvel_V1_abs ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2c) ## no group effect: X(1) = 0.0298 , p = 0.8629

```

# Create Multiplot for duration & amplitude

```{r, message = FALSE, warning = FALSE}

library(grid)
library(gridExtra)
library(cowplot)
library(lemon)

ggsave(filename = "../plots/TB_DUR_DISPL_CON_PD.png", 
       grid_arrange_shared_legend(plot01, plot02, ncol=1, nrow=2,
                    top = textGrob(" \nTongue Body Movement\n ",gp=gpar(fontsize=40,font="bold", margin(b = 20)))),
       width = 18, height = 18, dpi = 300, units = "in", device='png')

```

# Movement Profiles

Descriptive stats for reporting values in Table\

Subset for production of vowels /a/ and /i/

```{r, message = FALSE, warning = FALSE}

data_red_i_a <-  filter(data_red, vowel %in% c('i', 'a'))

```

Acceleration phase

```{r, message = FALSE, warning = FALSE}

ACC_stats <- data_red_i_a %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(acc_V1)),
            SD = round(sd(acc_V1)))
ACC_stats 

ACC_stats2 <- data_red_i_a %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(acc_V1)),
            SD = round(sd(acc_V1)))
ACC_stats2

```

Deceleration phase

```{r, message = FALSE, warning = FALSE}

DEC_stats <- data_red_i_a %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(dec_V1)),
            SD = round(sd(dec_V1)))
DEC_stats 

DEC_stats2 <- data_red_i_a %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(dec_V1)),
            SD = round(sd(dec_V1)))
DEC_stats2

```

Symmetry profile

```{r, message = FALSE, warning = FALSE}

SYMM_stats <- data_red_i_a %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(symmetry_V1),1),
            SD = round(sd(symmetry_V1),1))
SYMM_stats 

SYMM_stats2 <- data_red_i_a %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(symmetry_V1),1),
            SD = round(sd(symmetry_V1),1))
SYMM_stats2

```

Create linear mixed model to investigate disease effect on symmetry

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(symmetry_V1 ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red_i_a, REML = FALSE)

mdl2 <- lmer(symmetry_V1 ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red_i_a, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(symmetry_V1 ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red_i_a, REML = FALSE)

anova(mdl2, mdl2b) ## no effect of focus: X(2) = 2.3559, p = 0.3079

## check group

mdl2c <- lmer(symmetry_V1 ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red_i_a, REML = FALSE)

anova(mdl2, mdl2c) ## no group effect: X(1) = 0.0107, p = 0.9177

```

Stiffness

```{r, message = FALSE, warning = FALSE}

STIFF_stats <- data_red_i_a %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(stiffness_V1_abs),1),
            SD = round(sd(stiffness_V1_abs),1))
STIFF_stats 

STIFF_stats2 <- data_red_i_a %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(stiffness_V1_abs),1),
            SD = round(sd(stiffness_V1_abs),1))
STIFF_stats2

```

Create linear mixed model to investigate disease effect on stiffness

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(stiffness_V1_abs ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red_i_a, REML = FALSE)

mdl2 <- lmer(stiffness_V1_abs ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red_i_a, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(stiffness_V1_abs ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red_i_a, REML = FALSE)

anova(mdl2, mdl2b) ## no effect of focus: X(2) = 2.713, p = 0.2576

## check group

mdl2c <- lmer(stiffness_V1_abs ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red_i_a, REML = FALSE)

anova(mdl2, mdl2c) ## no group effect: X(1) = 2.2514, p =  0.1335

```

# Timing pattern

```{r, message = FALSE, warning = FALSE}

ONS_avgs <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(onsC1_to_onsV1)),
            SD = round(sd(onsC1_to_onsV1)))
ONS_avgs

TARG_avgs <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(targC1_to_targV1)),
            SD = round(sd(targC1_to_targV1)))
TARG_avgs

OnsAlign_avgs <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(onsV1.syll.dur),2)*100,
            SD = round(sd(onsV1.syll.dur),2)*100)
OnsAlign_avgs

TargAlign_avgs <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(targV1.syll.dur),2)*100,
            SD = round(sd(targV1.syll.dur),2)*100)
TargAlign_avgs

```

Statistical analysis for timing patterns

create model for onset to onset interval

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(onsC1_to_onsV1 ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

mdl2 <- lmer(onsC1_to_onsV1 ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(onsC1_to_onsV1 ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2b) ## no effect of focus: X(2) = 1.4016, p = 0.4962

## check group

mdl2c <- lmer(onsC1_to_onsV1 ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2c) ## no group effect: X(1) = 1.8463, p = 0.1742

```

create model for target to target interval

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(targC1_to_targV1 ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

mdl2 <- lmer(targC1_to_targV1 ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(targC1_to_targV1 ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2b) ## effect of focus: X(2) = 21.873, p = 1.779e-05

emmeans(mdl2, specs = pairwise ~ focus_con)

## check group

mdl2c <- lmer(targC1_to_targV1 ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2c) ## group effect: X(1) = 4.4016, p = 0.03591 *

emmeans(mdl2, specs = pairwise ~ med_con)

```

create model for onset alignment

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(onsV1.syll.dur ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

mdl2 <- lmer(onsV1.syll.dur ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(onsV1.syll.dur ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2b) ## no effect of focus:  X(2) = 4.3793, p = 0.112

## check group

mdl2c <- lmer(onsV1.syll.dur ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2c) ## no group effect:  X(1) = 0.3184, p = 0.5726

```

create model for target alignment

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(targV1.syll.dur ~ med_con * focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

mdl2 <- lmer(targV1.syll.dur ~ med_con + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus
mdl2b <- lmer(targV1.syll.dur ~ med_con + 1 + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2b) ## no effect of focus: X(2) = 0.2312, p = 0.8908

## check group

mdl2c <- lmer(targV1.syll.dur ~ 1 + focus_con + (1+focus_con|speaker) + (1|vowel) + (1|cons), data_red, REML = FALSE)

anova(mdl2, mdl2c) ## no group effect: X(1) = 0.6523, p = 0.4193


```
