---
title: "Articulatory Analysis"
author: "Tabea Thies"
date: "`r Sys.Date()`"
output: html_document
---

# Analysis of tongue body movements comparing med-OFF and med-ON

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load necessary packages

```{r load_tidyverse, warning = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(merTools)
library(sjPlot)
library(sjmisc)
```

Load data set

```{r load_counts, message = FALSE, warning = FALSE}
data <- read_csv('../data/articulation.csv')
```

Create subset for the comparison med-OFF vs med-ON

```{r, message = FALSE, warning = FALSE}
data_red <-  filter(data, med_con %in% c('med-OFF', 'med-ON'))
```

# Movement duration

Descriptive stats for reporting values in Section 7.2.3

```{r, message = FALSE, warning = FALSE}

DUR_stats <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(GAI_V1)),
            SD = round(sd(GAI_V1)))
DUR_stats 

DUR_stats2 <- data_red %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(GAI_V1)),
            SD = round(sd(GAI_V1)))
DUR_stats2

```
Plot tongue body movement duration

```{r, message = FALSE, warning = FALSE}

DUR_stats2 %>% 
  ggplot(aes(x=med_con, y=M, group = focus_con, col=focus_con)) + 
  geom_line(size = 2) +
  geom_point(size = 14)+
  theme_classic() +
  theme(legend.position = 'none')+
  theme(legend.title = element_blank(),
        legend.text = element_text(size = 36))+
  scale_color_manual(values = c("#8da0cb", "#66c2a5", "#fc8d62")) +
  xlab(NULL) +
  scale_x_discrete(labels = c('med-OFF','med-ON'))+
  coord_cartesian(ylim = c(172, 206)) + 
  scale_y_continuous(breaks = seq(180, 200, 10)) +
  ylab("Duration (ms)")+
  theme(axis.title.y = element_text(face = 'bold', size = 42, margin = margin(r = 30)))+
  theme(strip.text.x = element_text(size=38, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 36),
        axis.text.y = element_text(size = 36)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())
        
```


Create linear mixed model to investigate disease effect on TB duration

```{r, message = FALSE, warning = FALSE}

mdl <- lmer(GAI_V1 ~ med_con + focus_con + vowel + cons + age  + (1 + focus_con+med_con|speaker), data_red, REML = FALSE)

```

Look at model summary

```{r, message = FALSE, warning = FALSE}

summary(mdl)

```

extracting fit statistics for reporting in text

```{r, message = FALSE, warning = FALSE}

tab_model(mdl)

```

## Explore random effects with function REsim
Effects for individual levels are simulated\
Output: mean, median and sd of the random effect estimates\

Plots estimated random effects for each participant\
95 % Intervals that do not overlap zero are bold

```{r, message = FALSE, warning = FALSE}

randoms <- REsim(mdl) 
plotREsim(randoms)

```


Plot estimated 95 % interval spans for the intercept of TB duration

```{r, message = FALSE, warning = FALSE}

random2 <- subset(randoms, term == "(Intercept)")

random2 %>% 
  ggplot(aes(x = mean, xmin = mean-sd-sd, xmax = mean+sd+sd, y = reorder(groupID, mean))) +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  geom_point(size = 5)+
  geom_errorbar(size = 1) +
  theme_classic()+
  xlab('95 % intervals of estimated random intercepts for TB duration') +
  ylab('Speaker')+
  theme(axis.title.y = element_text(face = 'bold', size = 18, margin = margin(r = 12)))+
  theme(axis.title.x = element_text(face = 'bold', size = 18, margin = margin(t = 12)))+
  theme(strip.text.x = element_text(size=18, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())+
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 18))

```


Plot estimated 95 % interval spans for the slope of TB duration

```{r, message = FALSE, warning = FALSE}

random3 <- subset(randoms, term == "med_conmed-ON")

random3 %>% 
  ggplot(aes(x = mean, xmin = mean-sd-sd, xmax = mean+sd+sd, y = reorder(groupID, mean))) +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  geom_point(size = 5)+
  geom_errorbar(size = 1) +
  theme_classic()+
  xlab('95 % intervals of estimated random slopes for TB duration') +
  ylab('Speaker')+
  theme(axis.title.y = element_text(face = 'bold', size = 18, margin = margin(r = 12)))+
  theme(axis.title.x = element_text(face = 'bold', size = 18, margin = margin(t = 12)))+
  theme(strip.text.x = element_text(size=18, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())+
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 18))

```


# Movement amplitude

Descriptive stats for reporting values in Section 7.2.3

```{r, message = FALSE, warning = FALSE}

DISPL_stats <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(disp_V1_abs),1),
            SD = round(sd(disp_V1_abs),1))
DISPL_stats 

DISPL_stats2 <- data_red %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(disp_V1_abs),1),
            SD = round(sd(disp_V1_abs),1))
DISPL_stats2

```
Plot tongue body movement amplitude

```{r, message = FALSE, warning = FALSE}

DISPL_stats2 %>% 
  ggplot(aes(x=med_con, y=M, group = focus_con, col=focus_con)) + 
  geom_line(size = 2) +
  geom_point(size = 14)+
  theme_classic() +
  theme(legend.position = 'none')+
  theme(legend.title = element_blank(),
        legend.text = element_text(size = 36))+
  scale_color_manual(values = c("#8da0cb", "#66c2a5", "#fc8d62")) +
  coord_cartesian(ylim = c(7, 8.2)) + 
  scale_y_continuous(breaks = seq(7, 8, 0.5)) +
  xlab(NULL) +
  scale_x_discrete(labels = c('med-OFF','med-ON'))+
  ylab("Amplitude (mm)")+
  theme(axis.title.y = element_text(face = 'bold', size = 42, margin = margin(r = 30)))+
  theme(strip.text.x = element_text(size=38, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 36),
        axis.text.y = element_text(size = 36)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())
        
``` 


Create linear mixed model to investigate disease effect on TB amplitude

```{r, message = FALSE, warning = FALSE}

mdl2 <- lmer(disp_V1_abs ~ med_con + focus_con + vowel + cons + age  + (1 + focus_con+med_con|speaker), data_red, REML = FALSE)

```

Look at model summary

```{r, message = FALSE, warning = FALSE}

summary(mdl2)

```

extracting fit statistics for reporting in text

```{r, message = FALSE, warning = FALSE}

tab_model(mdl2)

```

## Explore random effects with function REsim
Effects for individual levels are simulated\
Output: mean, median and sd of the random effect estimates\

Plots estimated random effects for each participant\
95 % Intervals that do not overlap zero are bold

```{r, message = FALSE, warning = FALSE}

randoms <- REsim(mdl2) 
plotREsim(randoms)

```



Plot estimated 95 % interval spans for the intercept of TB amplitude

```{r, message = FALSE, warning = FALSE}

random2 <- subset(randoms, term == "(Intercept)")

random2 %>% 
  ggplot(aes(x = mean, xmin = mean-sd-sd, xmax = mean+sd+sd, y = reorder(groupID, mean))) +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  geom_point(size = 5)+
  geom_errorbar(size = 1) +
  theme_classic()+
  xlab('95 % intervals of estimated random intercepts for TB amplitude') +
  ylab('Speaker')+
  theme(axis.title.y = element_text(face = 'bold', size = 18, margin = margin(r = 12)))+
  theme(axis.title.x = element_text(face = 'bold', size = 18, margin = margin(t = 12)))+
  theme(strip.text.x = element_text(size=18, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())+
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 18))

```



Plot estimated 95 % interval spans for the slopes of TB amplitude

```{r, message = FALSE, warning = FALSE}

random3 <- subset(randoms, term == "med_conmed-ON")

random3 %>% 
  ggplot(aes(x = mean, xmin = mean-sd-sd, xmax = mean+sd+sd, y = reorder(groupID, mean))) +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  geom_point(size = 5)+
  geom_errorbar(size = 1) +
  theme_classic()+
  xlab('95 % intervals of estimated random slopes for TB amplitude') +
  ylab('Speaker')+
  theme(axis.title.y = element_text(face = 'bold', size = 18, margin = margin(r = 12)))+
  theme(axis.title.x = element_text(face = 'bold', size = 18, margin = margin(t = 12)))+
  theme(strip.text.x = element_text(size=18, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())+
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 18))

```


# Movement peak velocities

Descriptive stats for reporting values in Section 7.2.3

```{r, message = FALSE, warning = FALSE}

PVEL_stats <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(pvel_V1_abs)),
            SD = round(sd(pvel_V1_abs)))
PVEL_stats 

PVEL_stats2 <- data_red %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(pvel_V1_abs)),
            SD = round(sd(pvel_V1_abs)))
PVEL_stats2

```
Create linear mixed model to investigate disease effect on TB velocity

```{r, message = FALSE, warning = FALSE}

mdl3 <- lmer(pvel_V1_abs ~ med_con + focus_con + vowel + cons + age  + (1 + focus_con+med_con|speaker), data_red, REML = FALSE)

```

Look at model summary

```{r, message = FALSE, warning = FALSE}

summary(mdl3)

```

extracting fit statistics for reporting in text

```{r, message = FALSE, warning = FALSE}

tab_model(mdl3)

```

Explore random effects with function REsim


```{r, message = FALSE, warning = FALSE}
randoms <- REsim(mdl3) 
plotREsim(randoms)

```


Plot estimated 95 % interval spans for the intercept of TB velocity

```{r, message = FALSE, warning = FALSE}

random2 <- subset(randoms, term == "(Intercept)")

random2 %>% 
  ggplot(aes(x = mean, xmin = mean-sd-sd, xmax = mean+sd+sd, y = reorder(groupID, mean))) +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  geom_point(size = 5)+
  geom_errorbar(size = 1) +
  theme_classic()+
  xlab('95 % intervals of estimated random intercepts for TB velocity') +
  ylab('Speaker')+
  theme(axis.title.y = element_text(face = 'bold', size = 18, margin = margin(r = 12)))+
  theme(axis.title.x = element_text(face = 'bold', size = 18, margin = margin(t = 12)))+
  theme(strip.text.x = element_text(size=18, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())+
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 18))

```


Plot estimated 95 % interval spans for the slopes of TB velocity

```{r, message = FALSE, warning = FALSE}

random3 <- subset(randoms, term == "med_conmed-ON")

random3 %>% 
  ggplot(aes(x = mean, xmin = mean-sd-sd, xmax = mean+sd+sd, y = reorder(groupID, mean))) +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  geom_point(size = 5)+
  geom_errorbar(size = 1) +
  theme_classic()+
  xlab('95 % intervals of estimated random slopes for TB velocity') +
  ylab('Speaker')+
  theme(axis.title.y = element_text(face = 'bold', size = 18, margin = margin(r = 12)))+
  theme(axis.title.x = element_text(face = 'bold', size = 18, margin = margin(t = 12)))+
  theme(strip.text.x = element_text(size=18, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())+
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 18))

```


# Movement Profiles

Descriptive stats for reporting values in Table 7.2

Subset for production of vowels /a/ and /i/

```{r, message = FALSE, warning = FALSE}

data_red_i_a <-  filter(data_red, vowel %in% c('i', 'a'))

```

Acceleration phase

```{r, message = FALSE, warning = FALSE}

ACC_stats <- data_red_i_a %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(acc_V1)),
            SD = round(sd(acc_V1)))
ACC_stats 

ACC_stats2 <- data_red_i_a %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(acc_V1)),
            SD = round(sd(acc_V1)))
ACC_stats2

```

Deceleration phase

```{r, message = FALSE, warning = FALSE}

DEC_stats <- data_red_i_a %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(dec_V1)),
            SD = round(sd(dec_V1)))
DEC_stats 

DEC_stats2 <- data_red_i_a %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(dec_V1)),
            SD = round(sd(dec_V1)))
DEC_stats2

```

Symmetry profile

```{r, message = FALSE, warning = FALSE}

SYMM_stats <- data_red_i_a %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(symmetry_V1),1),
            SD = round(sd(symmetry_V1),1))
SYMM_stats 

SYMM_stats2 <- data_red_i_a %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(symmetry_V1),1),
            SD = round(sd(symmetry_V1),1))
SYMM_stats2

```

Stiffness

```{r, message = FALSE, warning = FALSE}

STIFF_stats <- data_red_i_a %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(stiffness_V1_abs),1),
            SD = round(sd(stiffness_V1_abs),1))
STIFF_stats 

STIFF_stats2 <- data_red_i_a %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(stiffness_V1_abs),1),
            SD = round(sd(stiffness_V1_abs),1))
STIFF_stats2

```

# Timing pattern

```{r, message = FALSE, warning = FALSE}

ONS_avgs <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(onsC1_to_onsV1)),
            SD = round(sd(onsC1_to_onsV1)))
ONS_avgs

TARG_avgs <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(targC1_to_targV1)),
            SD = round(sd(targC1_to_targV1)))
TARG_avgs

OnsAlign_avgs <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(onsV1.syll.dur),2)*100,
            SD = round(sd(onsV1.syll.dur),2)*100)
OnsAlign_avgs

TargAlign_avgs <- data_red %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(targV1.syll.dur),2)*100,
            SD = round(sd(targV1.syll.dur),2)*100)
TargAlign_avgs

```

Statistical analysis for timing patterns

create model for onset to onset interval

```{r, message = FALSE, warning = FALSE}

mdl4 <- lmer(onsC1_to_onsV1 ~ med_con + focus_con + vowel + cons + age  + (1 + focus_con+med_con|speaker), data_red, REML = FALSE)

```

Look at model

```{r, message = FALSE, warning = FALSE}

summary(mdl4)

```

Extracting fit statistics for reporting in text

```{r, message = FALSE, warning = FALSE}

tab_model(mdl4)

```

create model for target to target interval

```{r, message = FALSE, warning = FALSE}

mdl5 <- lmer(targC1_to_targV1 ~ med_con + focus_con + vowel + cons + age  + (1 + focus_con+med_con|speaker), data_red, REML = FALSE)

```

Look at model

```{r, message = FALSE, warning = FALSE}

summary(mdl5)

```

Extracting fit statistics for reporting in text

```{r, message = FALSE, warning = FALSE}

tab_model(mdl5)

```

create model for onset alignment

```{r, message = FALSE, warning = FALSE}

mdl6 <- lmer(onsV1.syll.dur ~ med_con + focus_con + vowel + cons + age  + (1 + focus_con+med_con|speaker), data_red, REML = FALSE)

```

Look at model

```{r, message = FALSE, warning = FALSE}

summary(mdl6)

```

Extracting fit statistics for reporting in text

```{r, message = FALSE, warning = FALSE}

tab_model(mdl6)

```

create model for target alignment

```{r, message = FALSE, warning = FALSE}

mdl7 <- lmer(targV1.syll.dur ~ med_con + focus_con + vowel + cons + age + (1 + focus_con+med_con|speaker), data_red, REML = FALSE)

```

Look at model

```{r, message = FALSE, warning = FALSE}

summary(mdl7)

```

Extracting fit statistics for reporting in text

```{r, message = FALSE, warning = FALSE}

tab_model(mdl7)

```