---
title: "Correlations"
author: "Tabea Thies"
date: "`r Sys.Date()`"
output: html_document
---

# Correlations between variabales

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load necessary packages

```{r load_tidyverse, warning = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
set.seed(42)

```

Load data set

```{r load_counts, message = FALSE, warning = FALSE}
acoustic <- read_csv('../data/acoustics.csv')
artic <- read_csv('../data/articulation.csv')
VAI <- read_csv('../data/formants_VAI.csv')
ratings <- read_csv('../data/listener_ratings.csv')
LEDD <- read_delim('../data/LEDD.csv')

```

Calculate means to reduce tables and to combine them

```{r, message = FALSE, warning = FALSE}

data_acoustic_avgs <-acoustic %>% 
  group_by(speaker, focus_con, med_con, targetword) %>% 
  summarise(Sylldur = mean(syll.dur, na.rm = T),
            V1dur = mean(V1.dur, na.rm = T))

data_artic_avgs <- artic %>% 
  group_by(speaker, focus_con, med_con, targetword) %>% 
  summarise(age = mean(age, na.rm = T),
            UPDRS = mean(UPDRS_total, na.rm = T),
            axial = mean(UPDRS_axial, na.rm = T),
            rigid = mean(UPDRS_rigidity, na.rm = T),
            brady = mean(UPDRS_brady, na.rm = T),
            PIGD = mean(UPDRS_PIGD, na.rm = T),
            
            VAS = mean(VAS, na.rm = T),
            BDI = mean(BDI, na.rm = T),
            PANDA = mean(PANDA, na.rm = T),
            MMST = mean(MMST, na.rm = T),
            
            GAI = mean(GAI_V1, na.rm = T),
            PVEL = mean(pvel_V1_abs, na.rm = T),
            DISP = mean(disp_V1_abs, na.rm = T),
            MMST = mean(MMST, na.rm = T))
```

Combine data frames

```{r, message = FALSE, warning = FALSE}

VAI$med_con <- recode(VAI$med_con, "CON" = "Control")

data_new1 <- inner_join(data_acoustic_avgs, VAI, by=c("speaker", "med_con", "focus_con"))
data_new2 <- merge(data_new1, data_artic_avgs, by=c("speaker", "med_con", "focus_con", "targetword", "age", "VAS", "BDI", "PANDA", "MMST"), all = T)

## reduce to only productions of /i, a, u/ which were elicited in Intell and Naturalness ratings

data_new3 <- filter(data_new2, targetword %in% c('lina', 'mila', 'lani', 'mali', 'luna', 'mula'))

data_new <- data_new3[ -c(12)]

```

Filter data only for broad productions to be able to match it with listener ratings

```{r, message = FALSE, warning = FALSE}

data_broad <- filter(data_new, focus_con %in% c('broad'))

colnames(ratings)[2] <- "targetword"
colnames(ratings)[3] <- "med_con"
colnames(ratings)[4] <- "speaker"

ratings$med_con <- recode(ratings$med_con,
                          "CON" = "Control",
                          "medOFF" = "med-OFF",
                          "medON" = "med-ON",
                          "DBSOFF" = "med-OFF/DBS-OFF",
                          "DBSON" = "med-OFF/DBS-ON")

ratings_avgs <- ratings %>% 
  group_by(speaker, med_con, targetword) %>% 
  summarise(Intell = mean(Intell, na.rm = T),
            Natur = mean(Natural, na.rm = T))

data_broad_new <- merge(data_broad, ratings_avgs, by=c("speaker", "med_con", "targetword"), all = T)

```

Subsets per condition

```{r, message = FALSE, warning = FALSE}

data_broad_CON <- filter(data_broad_new, med_con %in% c('Control'))
data_broad_medOFF <- filter(data_broad_new, med_con %in% c('med-OFF'))
data_broad_medON <- filter(data_broad_new, med_con %in% c('med-ON'))
data_broad_DBSON <- filter(data_broad_new, med_con %in% c('med-OFF/DBS-ON'))

```

# Statistical analysis: What explains reduced intelligibility?

Acoustic vowel duration ~ Intelligibility

```{r, message = FALSE, warning = FALSE}

ggplot(data_broad_new, aes(x=V1dur, y=Intell)) + 
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(.~med_con)

```

Calculate correlation coefficient between V1.dur and Intell
 
```{r, message = FALSE, warning = FALSE}

## Controls
shapiro.test(data_broad_CON$Intell) # p = 6.595e-05
shapiro.test(data_broad_CON$V1dur) # p =  0.1047

cor.test(data_broad_CON$Intell, data_broad_CON$V1dur, method = "spearman") ## r(75) = -0.46, p = 2.816e-05

## med-OFF
shapiro.test(data_broad_medOFF$Intell) # p = 0.05947
shapiro.test(data_broad_medOFF$V1dur) # p = 0.005831

cor.test(data_broad_medOFF$Intell, data_broad_medOFF$V1dur, method = "spearman") ## r(74) = 0.10, p = 0.3609

## med-ON
shapiro.test(data_broad_medON$Intell) # p = 0.05628
shapiro.test(data_broad_medON$V1dur) # p = 0.1749

cor.test(data_broad_medON$Intell, data_broad_medON$V1dur, method = "pearson") ## r(74) = 0.12, p = 0.2938

## DBS-ON
shapiro.test(data_broad_DBSON$Intell) # p = 0.0002829
shapiro.test(data_broad_DBSON$V1dur) # p = 0.08312

cor.test(data_broad_DBSON$Intell, data_broad_DBSON$V1dur, method = "spearman") ## r(75) = 0.08, p = 0.4885
```

VAI ~ Intelligibility

```{r, message = FALSE, warning = FALSE}

ggplot(data_broad_new, aes(x=VAI_meas, y=Intell)) + 
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(.~med_con)

```

Calculate correlation coefficient between VAI and Intell
 
```{r, message = FALSE, warning = FALSE}

## Controls
shapiro.test(data_broad_CON$Intell) # p = 6.595e-05
shapiro.test(data_broad_CON$VAI_meas) # p = 0.001078

cor.test(data_broad_CON$Intell, data_broad_CON$VAI_meas, method = "spearman") ## r(75) = 0.25, p = 0.02931

## med-OFF
shapiro.test(data_broad_medOFF$Intell) # p =  0.05947
shapiro.test(data_broad_medOFF$VAI_meas) # p = 9.203e-06

cor.test(data_broad_medOFF$Intell, data_broad_medOFF$VAI_meas, method = "spearman") ## r(74) = 0.43 p = 0.000132

## med-ON
shapiro.test(data_broad_medON$Intell) # p = 0.05628
shapiro.test(data_broad_medON$VAI_meas) # p = 0.004046

cor.test(data_broad_medON$Intell, data_broad_medON$VAI_meas, method = "spearman") ## r(74) = -0.04, n.s.

## DBS-ON
shapiro.test(data_broad_DBSON$Intell) # p = 0.0002829
shapiro.test(data_broad_DBSON$VAI_meas) # p = 0.0005819

cor.test(data_broad_DBSON$Intell, data_broad_DBSON$VAI_meas, method = "spearman") ## r(75) = 0.50, p =  3.715e-06
```

Tongue Body Movement Duration ~ Intelligibility

```{r, message = FALSE, warning = FALSE}

ggplot(data_broad_new, aes(x=GAI, y=Intell)) + 
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(.~med_con)

```

Calculate correlation coefficient between TB duration and Intell
 
```{r, message = FALSE, warning = FALSE}

## Controls
shapiro.test(data_broad_CON$Intell) # p = 6.595e-05
shapiro.test(data_broad_CON$GAI) # p =  0.1592

cor.test(data_broad_CON$Intell, data_broad_CON$GAI, method = "spearman") ## r(75) = -0.30, p = 0.007766

## med-OFF
shapiro.test(data_broad_medOFF$Intell) # p = 0.05947
shapiro.test(data_broad_medOFF$GAI) # p = 0.7077

cor.test(data_broad_medOFF$Intell, data_broad_medOFF$GAI, method = "pearson") ## r(74) = 0.06, n.s.

## med-ON
shapiro.test(data_broad_medON$Intell) # p =  0.05628
shapiro.test(data_broad_medON$GAI) # p = 0.02441

cor.test(data_broad_medON$Intell, data_broad_medON$GAI, method = "spearman") ## r(74) = 0.10, n.s.

## DBS-ON
shapiro.test(data_broad_DBSON$Intell) # p = 0.0002829
shapiro.test(data_broad_DBSON$GAI) # p = 0.0002742

cor.test(data_broad_DBSON$Intell, data_broad_DBSON$GAI, method = "spearman") ## r(75) = 0.18, n.s.
```

Tongue Body Movement Amplitude ~ Intelligibility

```{r, message = FALSE, warning = FALSE}

ggplot(data_broad_new, aes(x=DISP, y=Intell)) + 
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(.~med_con)

```

# Statistical analysis: What explains reduced naturalness?

Acoustic vowel duration ~ Naturalness

```{r, message = FALSE, warning = FALSE}

ggplot(data_broad_new, aes(x=V1dur, y=Natur)) + 
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(.~med_con)

```

Calculate correlation coefficient between V1.dur and Natur
 
```{r, message = FALSE, warning = FALSE}

## Controls
shapiro.test(data_broad_CON$Natur) # p =  0.4031
shapiro.test(data_broad_CON$V1dur) # p = 0.1047

cor.test(data_broad_CON$Natur, data_broad_CON$V1dur, method = "pearson") ## r(75) = -0.29, p = .011

## med-OFF
shapiro.test(data_broad_medOFF$Natur) # p = 4.867e-07
shapiro.test(data_broad_medOFF$V1dur) # p = 0.005831

cor.test(data_broad_medOFF$Natur, data_broad_medOFF$V1dur, method = "spearman") ## r(74) = -0.23, p = .044.

## med-ON
shapiro.test(data_broad_medON$Natur) # p = 0.0002095
shapiro.test(data_broad_medON$V1dur) # p = 0.1749

cor.test(data_broad_medON$Natur, data_broad_medON$V1dur, method = "spearman") ## r(74) = -0.42, p = 0.0001967

## DBS-ON
shapiro.test(data_broad_DBSON$Natur) # p = 0.0001987
shapiro.test(data_broad_DBSON$V1dur) # p = 0.08312

cor.test(data_broad_DBSON$Natur, data_broad_DBSON$V1dur, method = "spearman") ## r(75) = -0.40, p = 0.0004194
```

VAI ~ Naturalness

```{r, message = FALSE, warning = FALSE}

ggplot(data_broad_new, aes(x=VAI_meas, y=Natur)) + 
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(.~med_con)

```

Calculate correlation coefficient between VAI and Natur
 
```{r, message = FALSE, warning = FALSE}

## Controls
shapiro.test(data_broad_CON$Natur) # p = 0.4031
shapiro.test(data_broad_CON$VAI_meas) # p = 0.001078

cor.test(data_broad_CON$Natur, data_broad_CON$VAI_meas, method = "spearman") ## r(75) = 0.11, n.s.

## med-OFF
shapiro.test(data_broad_medOFF$Natur) # p = 4.867e-07
shapiro.test(data_broad_medOFF$VAI_meas) # p = 9.203e-06

cor.test(data_broad_medOFF$Natur, data_broad_medOFF$VAI_meas, method = "spearman") ## r(74) = 0.59, p = 1.635e-08

## med-ON
shapiro.test(data_broad_medON$Natur) # p = 0.0002095
shapiro.test(data_broad_medON$VAI_meas) # p = 0.004046

cor.test(data_broad_medON$Natur, data_broad_medON$VAI_meas, method = "spearman") ## r(74) = 0.25, p = 0.02727

## DBS-ON
shapiro.test(data_broad_DBSON$Natur) # p = 0.0001987
shapiro.test(data_broad_DBSON$VAI_meas) # p = 0.0005819

cor.test(data_broad_DBSON$Natur, data_broad_DBSON$VAI_meas, method = "spearman") ## r(75) = 0.47, p = 1.337e-05
```

Tongue Body Movement Duration ~ Naturalness

```{r, message = FALSE, warning = FALSE}

ggplot(data_broad_new, aes(x=GAI, y=Natur)) + 
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(.~med_con)

```

Calculate correlation coefficient between TB duration and Natur
 
```{r, message = FALSE, warning = FALSE}

## Controls
shapiro.test(data_broad_CON$Natur) # p = 0.4031
shapiro.test(data_broad_CON$GAI) # p = 0.1592

cor.test(data_broad_CON$Natur, data_broad_CON$GAI, method = "pearson") ## r(75) = -0.12, n.s.

## med-OFF
shapiro.test(data_broad_medOFF$Natur) # p = 4.867e-07
shapiro.test(data_broad_medOFF$GAI) # p =  0.7077

cor.test(data_broad_medOFF$Natur, data_broad_medOFF$GAI, method = "spearman") ## r(74) = -0.11, n.s.

## med-ON
shapiro.test(data_broad_medON$Natur) # p = 0.0002095
shapiro.test(data_broad_medON$GAI) # p = 0.02441

cor.test(data_broad_medON$Natur, data_broad_medON$GAI, method = "spearman") ## r(74) = -0.38, p = 0.0008144

## DBS-ON
shapiro.test(data_broad_DBSON$Natur) # p = 0.0001987
shapiro.test(data_broad_DBSON$GAI) # p = 0.0002742

cor.test(data_broad_DBSON$Natur, data_broad_DBSON$GAI, method = "spearman") ## r(75) = -0.22, n.s
```

# Correlation between other parameters:

all focus conditions are considered

Subsets per group/ treatment condition

```{r, message = FALSE, warning = FALSE}

data_CON <- filter(data_new, med_con %in% c('Control'))
data_medOFF <- filter(data_new, med_con %in% c('med-OFF'))
data_medON <- filter(data_new, med_con %in% c('med-ON'))
data_DBSON <- filter(data_new, med_con %in% c('med-OFF/DBS-ON'))

```

# Which factors are responsible for longer acoustic durations

Acoustic vowel duration ~ TB DUR

```{r, message = FALSE, warning = FALSE}

ggplot(data_new, aes(x=V1dur, y=GAI)) + 
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(.~med_con)

```

Calculate correlation coefficient between V1.dur and TB DUR
 
```{r, message = FALSE, warning = FALSE}

## Control
shapiro.test(data_CON$GAI) # p = 0.001196
shapiro.test(data_CON$V1dur) # p = 0.0008486

cor.test(data_CON$GAI, data_CON$V1dur, method = "spearman") ## r(232) = 0.57, p < .001

## med-OFF
shapiro.test(data_medOFF$GAI) # p =  0.00677
shapiro.test(data_medOFF$V1dur) # p = 2.811e-05

cor.test(data_medOFF$GAI, data_medOFF$V1dur, method = "spearman") ## r(229) = 0.55, p < .001

## med-ON
shapiro.test(data_medON$GAI) # p = 3.426e-08
shapiro.test(data_medON$V1dur) # p = 0.01099

cor.test(data_medON$GAI, data_medON$V1dur, method = "spearman") ## r(228) = 0.60, p < .001

## DBS-ON
shapiro.test(data_DBSON$GAI) # p = 4.34e-06
shapiro.test(data_DBSON$V1dur) # p = 0.0008349

cor.test(data_DBSON$GAI, data_DBSON$V1dur, method = "spearman") ## r(214) = 0.62, p < .001
```

# Which factors are responsible for smaller VAI values?

VAI ~ UPDRS

```{r, message = FALSE, warning = FALSE}

ggplot(data_new, aes(x=VAI_meas, y=UPDRS)) + 
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(.~med_con)

```

Calculate correlation coefficient between VAI and UPDRS
 
```{r, message = FALSE, warning = FALSE}


## med-OFF
shapiro.test(data_medOFF$VAI_meas) # p =  2.41e-09
shapiro.test(data_medOFF$UPDRS) # p = 1.239e-09

cor.test(data_medOFF$VAI_meas, data_medOFF$UPDRS, method = "spearman") ## r(229) = -0.39, p = 7.971e-10

## med-ON
shapiro.test(data_medON$VAI_meas) # p = 0.0007983
shapiro.test(data_medON$UPDRS) # p = 7.083e-11

cor.test(data_medON$VAI_meas, data_medON$UPDRS, method = "spearman") ## r(228) = -0.25, p = 0.0001303

## DBS-ON
shapiro.test(data_DBSON$VAI_meas) # p = 0.001383
shapiro.test(data_DBSON$UPDRS) # p = 1.003e-09

cor.test(data_DBSON$VAI_meas, data_DBSON$UPDRS, method = "spearman") ## r(214) = -0.23, p = 0.0006561
```

VAI ~ axial symptoms

```{r, message = FALSE, warning = FALSE}

ggplot(data_new, aes(x=VAI_meas, y=axial)) + 
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(.~med_con)

```

Calculate correlation coefficient between VAI and axial subscores
 
```{r, message = FALSE, warning = FALSE}


## med-OFF
shapiro.test(data_medOFF$VAI_meas) # p = 2.41e-09
shapiro.test(data_medOFF$axial) # p < 2.2e-16

cor.test(data_medOFF$VAI_meas, data_medOFF$axial, method = "spearman") ## r(229) = -0.59, p < 2.2e-16

## med-ON
shapiro.test(data_medON$VAI_meas) # p = 0.0007983
shapiro.test(data_medON$axial) # p < 2.2e-16

cor.test(data_medON$VAI_meas, data_medON$axial, method = "spearman") ## r(228) = -0.15, p = 0.02551

## DBS-ON
shapiro.test(data_DBSON$VAI_meas) # p = 0.001383
shapiro.test(data_DBSON$axial) # p = 1.094e-10

cor.test(data_DBSON$VAI_meas, data_DBSON$axial, method = "pearson") ## r(214) = -0.30, p = 7.573e-06
```

VAI ~ bradykinesia symptoms

```{r, message = FALSE, warning = FALSE}

ggplot(data_new, aes(x=VAI_meas, y=brady)) + 
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(.~med_con)

```

Calculate correlation coefficient between VAI and brady subscores
 
```{r, message = FALSE, warning = FALSE}


## med-OFF
shapiro.test(data_medOFF$VAI_meas) # p = 2.41e-09
shapiro.test(data_medOFF$brady) # p = 4.862e-11

cor.test(data_medOFF$VAI_meas, data_medOFF$brady, method = "spearman") ## r(229) = -0.39, p = 1.137e-09

## med-ON
shapiro.test(data_medON$VAI_meas) # p = 0.0007983
shapiro.test(data_medON$brady) # p = 7.509e-12

cor.test(data_medON$VAI_meas, data_medON$brady, method = "spearman") ## r(228) = -0.03, n.s.

## DBS-ON
shapiro.test(data_DBSON$VAI_meas) # p = 0.001383
shapiro.test(data_DBSON$brady) # p = 3.908e-13


cor.test(data_DBSON$VAI_meas, data_DBSON$brady, method = "spearman") ## r(214) = -0.15, p = 0.0315
```

VAI ~ symptoms of Posture and Gait

```{r, message = FALSE, warning = FALSE}

ggplot(data_new, aes(x=VAI_meas, y=PIGD)) + 
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(.~med_con)

```

Calculate correlation coefficient between VAI and PIGD
 
```{r, message = FALSE, warning = FALSE}

## med-OFF
shapiro.test(data_medOFF$VAI_meas) # p = 2.41e-09
shapiro.test(data_medOFF$PIGD) # p < 2.2e-16

cor.test(data_medOFF$VAI_meas, data_medOFF$PIGD, method = "spearman") ## r(229) = -0.54, p < 2.2e-16

## med-ON
shapiro.test(data_medON$VAI_meas) # p = 0.0007983
shapiro.test(data_medON$PIGD) # p =  2.2e-16

cor.test(data_medON$VAI_meas, data_medON$PIGD, method = "spearman") ## r(228) = -0.21, p =  0.001342

## DBS-ON
shapiro.test(data_DBSON$VAI_meas) # p = 0.001383
shapiro.test(data_DBSON$PIGD) # p = 6.968e-12

cor.test(data_DBSON$VAI_meas, data_DBSON$PIGD, method = "spearman") ## r(214) = -0.32, p =  1.261e-06
```

## What parameters affect self-perceived speaking ability of patients?

Acoustic vowel duration ~ VAS

```{r, message = FALSE, warning = FALSE}

ggplot(data_new, aes(x=V1dur, y=VAS)) + 
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(.~med_con)

```

Calculate correlation coefficient between V1.dur and VAS
 
```{r, message = FALSE, warning = FALSE}

## med-OFF
shapiro.test(data_medOFF$V1dur) # p = 2.811e-05
shapiro.test(data_medOFF$VAS) # p = 2.287e-10

cor.test(data_medOFF$V1dur, data_medOFF$VAS, method = "spearman") ## r(229) = 0.33, p = 1.947e-07

## med-ON
shapiro.test(data_medON$V1dur) # p = 0.01099
shapiro.test(data_medON$VAS) # p = 5.113e-15

cor.test(data_medON$V1dur, data_medON$VAS, method = "spearman") ## r(228) = 0.28, p = 1.393e-05

## DBS-ON
shapiro.test(data_DBSON$V1dur) # p = 0.0008928
shapiro.test(data_DBSON$VAS) # p = 5.686e-14

cor.test(data_DBSON$V1dur, data_DBSON$VAS, method = "spearman") ## r(214) = 0.34, p = 1.197e-07
```

VAI ~ VAS

```{r, message = FALSE, warning = FALSE}

ggplot(data_new, aes(x=VAI_meas, y=VAS)) + 
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(.~med_con)

```

Calculate correlation coefficient between VAI and VAS
 
```{r, message = FALSE, warning = FALSE}

## med-OFF
shapiro.test(data_medOFF$VAI_meas) # p = 2.41e-09
shapiro.test(data_medOFF$VAS) # p = 2.287e-10

cor.test(data_medOFF$VAI_meas, data_medOFF$VAS, method = "spearman") ## r(229) = 0.07, n.s.

## med-ON
shapiro.test(data_medON$VAI_meas) # p = 0.0007983
shapiro.test(data_medON$VAS) # p = 5.113e-15

cor.test(data_medON$VAI_meas, data_medON$VAS, method = "spearman") ## r(228) = 0.16, p = 0.0163

## DBS-ON
shapiro.test(data_DBSON$VAI_meas) # p =  0.001137
shapiro.test(data_DBSON$VAS) # p = 5.686e-14

cor.test(data_DBSON$VAI_meas, data_DBSON$VAS, method = "spearman") ## r(214) = 0.37, p = 7.265e-09
```

TB DUR ~ VAS

```{r, message = FALSE, warning = FALSE}

ggplot(data_new, aes(x=GAI, y=VAS)) + 
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(.~med_con)

```

Calculate correlation coefficient between TB DUR and VAS
 
```{r, message = FALSE, warning = FALSE}

## med-OFF
shapiro.test(data_medOFF$GAI) # p = 0.00677
shapiro.test(data_medOFF$VAS) # p = 2.287e-10

cor.test(data_medOFF$GAI, data_medOFF$VAS, method = "spearman") ## r(229) = 0.12, n.s.

## med-ON
shapiro.test(data_medON$GAI) # p = 3.426e-08
shapiro.test(data_medON$VAS) # p = 5.113e-15

cor.test(data_medON$GAI, data_medON$VAS, method = "spearman") ## r(228) = 0.17, p = 0.009093

## DBS-ON
shapiro.test(data_DBSON$GAI) # p =  4.34e-06
shapiro.test(data_DBSON$VAS) # p = 5.686e-14

cor.test(data_DBSON$GAI, data_DBSON$VAS, method = "spearman") ## r(214) = 0.32, p = 4.758e-07
```

UPDRS ~ VAS

```{r, message = FALSE, warning = FALSE}

ggplot(data_broad_new, aes(x=UPDRS, y=VAS)) + 
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(.~med_con)

```

Calculate correlation coefficient between UPDRS and VAS
 
```{r, message = FALSE, warning = FALSE}


## med-OFF
shapiro.test(data_medOFF$UPDRS) # p = 1.239e-09
shapiro.test(data_medOFF$VAS) # p = 2.287e-10

cor.test(data_broad_medOFF$UPDRS, data_broad_medOFF$VAS, method = "spearman") ## r(229) = -0.16, p = 0.01735

## med-ON
shapiro.test(data_medON$UPDRS) # p = 7.083e-11
shapiro.test(data_medON$VAS) # p = 5.113e-15

cor.test(data_broad_medON$UPDRS, data_broad_medON$VAS, method = "spearman") ## r(228) = -0.17, n.s.

## DBS-ON
shapiro.test(data_DBSON$UPDRS) # p = 1.003e-09
shapiro.test(data_DBSON$VAS) # p =  1.379e-13

cor.test(data_broad_DBSON$UPDRS, data_broad_DBSON$VAS, method = "spearman") ## r(76) = -0.58, p < 3.293e-08
```

axial ~ VAS

```{r, message = FALSE, warning = FALSE}

ggplot(data_new, aes(x=axial, y=VAS)) + 
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(.~med_con)

```

Calculate correlation coefficient between axial subscores and VAS
 
```{r, message = FALSE, warning = FALSE}


## med-OFF
shapiro.test(data_medOFF$axial) # p = 1.661e-07
shapiro.test(data_medOFF$VAS) # p = 0.008463

cor.test(data_broad_medOFF$axial, data_broad_medOFF$VAS, method = "spearman") ## r(75) = -0.05, n.s.

## med-ON
shapiro.test(data_medON$axial) # p = 9.958e-08
shapiro.test(data_medON$VAS) # p = 5.054e-05

cor.test(data_broad_medON$axial, data_broad_medON$VAS, method = "spearman") ## r(74) = -0.28, p = 0.01336

## DBS-ON
shapiro.test(data_DBSON$axial) # p = 0.003841
shapiro.test(data_DBSON$VAS) # p = 0.0002832

cor.test(data_broad_DBSON$axial, data_broad_DBSON$VAS, method = "spearman") ## r(76) = -0.43, p =  7.706e-05
```

## Correlations OFF to OFF change

Reported under section 10.2.4.

Combine data frames and subset for relevant conditions

```{r, message = FALSE, warning = FALSE}

data_PD <- filter(data_new, med_con != c('Control'))
data_PD <- merge(data_PD, LEDD, by="speaker", all = T)

data_medOFF <-  filter(data_PD, med_con %in% c('med-OFF'))
data_DBSOFF <-  filter(data_PD, med_con %in% c('med-OFF/DBS-OFF'))

```

Calculate means per condition

```{r, message = FALSE, warning = FALSE}

medOFF_avgs <- data_medOFF %>% 
  group_by(speaker) %>% 
  summarise(UPDRS_medOFF = mean(UPDRS, na.rm = T),
            axial_medOFF = mean(axial, na.rm = T),
            rigid_medOFF = mean(rigid, na.rm = T),
            brady_medOFF = mean(brady, na.rm = T),
            PIGD_medOFF = mean(PIGD, na.rm = T),
            
            GAI_medOFF = mean(GAI, na.rm = T),
            PVEL_medOFF = mean(PVEL, na.rm = T),
            DISP_medOFF = mean(DISP, na.rm = T),

            LEDD_medOFF = mean(LEDD_preop, na.rm = T),
            agonists_medOFF = mean(Agonists_preop, na.rm = T),
            time_agonist_medOFF = mean(last_agonist_preop, na.rm = T),
            time_visits = mean(Time_preop_postop))

DBSOFF_avgs <- data_DBSOFF  %>% 
  group_by(speaker) %>% 
  summarise(UPDRS_DBSOFF = mean(UPDRS, na.rm = T),
            axial_DBSOFF = mean(axial, na.rm = T),
            rigid_DBSOFF = mean(rigid, na.rm = T),
            brady_DBSOFF = mean(brady, na.rm = T),
            PIGD_DBSOFF = mean(PIGD, na.rm = T),
            
            GAI_DBSOFF = mean(GAI, na.rm = T),
            PVEL_DBSOFF = mean(PVEL, na.rm = T),
            DISP_DBSOFF = mean(DISP, na.rm = T),
            
            LEDD_DBSOFF = mean(LEDD_postop, na.rm = T),
            agonists_DBSOFF = mean(Agonists_postop, na.rm = T),
            time_agonist_DBSOFF = mean(last_agonist_postop, na.rm = T))

data_change <- inner_join(medOFF_avgs, DBSOFF_avgs, by= c("speaker"))

```

Calculate absolute and percentage change

```{r, message = FALSE, warning = FALSE}

data_change$UPDRS_abs <- (data_change$UPDRS_medOFF - data_change$UPDRS_DBSOFF) * -1
data_change$UPDRS_perc <- (data_change$UPDRS_medOFF - data_change$UPDRS_DBSOFF)/data_change$UPDRS_medOFF * 100

data_change$axial_abs <- (data_change$axial_medOFF - data_change$axial_DBSOFF) * -1
data_change$axial_perc <- (data_change$axial_medOFF - data_change$axial_DBSOFF)/data_change$axial_medOFF * 100

data_change$brady_abs <- (data_change$brady_medOFF - data_change$brady_DBSOFF) * -1
data_change$brady_perc <- (data_change$brady_medOFF - data_change$brady_DBSOFF)/data_change$brady_medOFF * 100

data_change$PIGD_abs <- (data_change$PIGD_medOFF - data_change$PIGD_DBSOFF) * -1
data_change$PIGD_perc <- (data_change$PIGD_medOFF - data_change$PIGD_DBSOFF)/data_change$PIGD_medOFF * 100

data_change$DISP_abs <- (data_change$DISP_medOFF - data_change$DISP_DBSOFF) * -1
data_change$DISP_perc <- (data_change$DISP_medOFF - data_change$DISP_DBSOFF)/data_change$DISP_medOFF * 100

data_change$PVEL_abs <- (data_change$PVEL_medOFF - data_change$PVEL_DBSOFF) * -1
data_change$PVEL_perc <- (data_change$PVEL_medOFF - data_change$PVEL_DBSOFF)/data_change$PVEL_medOFF * 100

data_change$LEDD_abs <- (data_change$LEDD_medOFF - data_change$LEDD_DBSOFF) * -1
data_change$LEDD_perc <- (data_change$LEDD_medOFF - data_change$LEDD_DBSOFF)/data_change$LEDD_medOFF * 100

data_change$agonists_abs <- (data_change$agonists_medOFF - data_change$agonists_DBSOFF) * -1
data_change$agonists_perc <- (data_change$agonists_medOFF - data_change$agonists_DBSOFF)/data_change$agonists_medOFF * 100

```

Correlations

Time and change in UPDRS

```{r, message = FALSE, warning = FALSE}

plot01 <- ggplot(data_change, aes(x=time_visits, y=UPDRS_abs))+
  geom_hline(yintercept=0, linetype="dashed", color = "gray", size=1)+
  geom_point(size = 5)+
  geom_smooth(method=lm, se=FALSE, size = 2, color = "black")+
  geom_text(aes(label = speaker), vjust = -.9, size = 6)+
  annotate("text", x = 14, y = 16, label = "r(10) = .53, p > .05", size = 9, color = "darkgray")+
  theme_classic() +
  ggtitle("")+
  ylab("Absolute change of UPDRS scores")+
  xlab("Time (in months)")+
  ylim(-12,20)+
  theme(plot.title = element_text(size = 40)) +
  theme(axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())+
  theme(axis.title.y = element_text(face = 'bold', size = 18, margin = margin(r = 12)))+
  theme(axis.title.x = element_text(face = 'bold', size = 18, margin = margin(t = 20)))

ggsave(plot01, filename = '../plots/UPDRS_OFF_time_preop_postop.png',
       width = 10, height = 5, dpi = 300)

plot02 <- ggplot(data_change, aes(x=time_visits, y=UPDRS_perc))+
  geom_hline(yintercept=0, linetype="dashed", color = "gray", size=1)+
  geom_point(size = 5)+
  geom_smooth(method=lm, se=FALSE, size = 2, color = "black")+
  geom_text(aes(label = speaker), vjust = -.9, size = 6)+
  annotate("text", x = 16, y = -80, label = "r(10) = .53, p > .05", size = 9, color = "darkgray")+
  theme_classic() +
  ggtitle("")+
  ylab("Percentage change of\nUPDRS III scores")+
  xlab("Time (in months)")+
  ylim(-110,45)+
  theme(plot.title = element_text(size = 40)) +
  theme(axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())+
  theme(axis.title.y = element_text(face = 'bold', size = 18, margin = margin(r = 12)))+
  theme(axis.title.x = element_text(face = 'bold', size = 18, margin = margin(t = 20)))

ggsave(plot02, filename = '../plots/UPDRS_OFF_percentage_time_preop_postop.png',
       width = 10, height = 5, dpi = 300)

shapiro.test(data_change$time_visits) # p = 0.06089
shapiro.test(data_change$UPDRS_abs) # p = 0.4317
shapiro.test(data_change$UPDRS_perc) # p = 0.1191

## TIME since first visit and absolute UPDRS change 
cor.test(data_change$time_visits, data_change$UPDRS_abs, method = "pearson") #  r(10) = -0.58, p-value = 0.04744

## TIME since first visit and absolute UPDRS change 
cor.test(data_change$time_visits, data_change$UPDRS_perc, method = "pearson") #  r(10) = 0.53, p-value = 0.07442

```

Change in AGONISTS ~ CHANGE in TB AMPLITUDE

```{r, message = FALSE, warning = FALSE}

ggplot(data_change, aes(x=agonists_perc, y=DISP_perc)) + 
  geom_point()+
  geom_smooth(method=lm)

shapiro.test(data_change$agonists_perc) # p = 0.1761
shapiro.test(data_change$DISP_perc) # p = 0.6581

cor.test(data_change$DISP_perc, data_change$agonists_perc, method = "pearson") ## r(10) = -0.39, p = 0.2059

```

Change in UPDRS ~ CHANGE in TB AMPLITUDE

```{r, message = FALSE, warning = FALSE}

ggplot(data_change, aes(x=UPDRS_perc, y=DISP_perc)) + 
  geom_point()+
  geom_smooth(method=lm)+
    geom_text(aes(label = speaker), vjust = -.9, size = 6)

shapiro.test(data_change$UPDRS_perc) # p = 0.1191
shapiro.test(data_change$DISP_perc) # p = 0.6581

cor.test(data_change$DISP_perc, data_change$UPDRS_perc, method = "pearson") ## r = 0.34, n.s.

```

Change in axial ~ CHANGE in TB AMPLITUDE

```{r, message = FALSE, warning = FALSE}

ggplot(data_change, aes(x=axial_perc, y=DISP_perc)) + 
  geom_point()+
  geom_smooth(method=lm)

shapiro.test(data_change$axial_perc) # p = 0.001338
shapiro.test(data_change$DISP_perc) # p = 0.6581

cor.test(data_change$DISP_perc, data_change$axial_perc, method = "spearman") ## r = -0.18, n.s.

```
