---
title: "Drug Dosages & Time between vistis"
author: "Tabea Thies"
date: "`r Sys.Date()`"
output: html_document
---

# Analysis of levodopa equivalent daily dose

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load necessary packages

```{r load_tidyverse, warning = FALSE, message = FALSE}
library(tidyverse)
```

Load data set

```{r load_counts, message = FALSE, warning = FALSE}
data <- read_delim('../data/LEDD.csv')
```

Calculate the means
Time (in months) between date of surgery and preoperative assessment: reported under section 8.1

```{r, message = FALSE, warning = FALSE}

Time_OP <- data %>% 
  summarise(M = round(mean(Time_OP_Test, na.rm = T)),
            SD = round(sd(Time_OP_Test, na.rm = T)))
Time_OP

```

Calculate the means
Time (in months) between preoperative visit and postoperative visits: reported under section 10.1

```{r, message = FALSE, warning = FALSE}

Time_preop_postop <- data %>% 
  summarise(M = round(mean(Time_preop_postop, na.rm = T)),
            SD = round(sd(Time_preop_postop, na.rm = T)))
Time_preop_postop 

```

Calculate means of levodopa equivalent daily doses (LEDD): reported under section 10.1

```{r, message = FALSE, warning = FALSE}

LEDD_preop <- data %>% 
  summarise(M = round(mean(LEDD_preop, na.rm = T)),
            SD = round(sd(LEDD_preop, na.rm = T)))
LEDD_preop

LEDD_postop <- data %>% 
  summarise(M = round(mean(LEDD_postop, na.rm = T)),
            SD = round(sd(LEDD_postop, na.rm = T)))
LEDD_postop

Agonist_preop <- data %>% 
  summarise(M = round(mean(Agonists_preop, na.rm = T)),
            SD = round(sd(Agonists_preop, na.rm = T)))
Agonist_preop

Agonist_postop <- data %>% 
  summarise(M = round(mean(Agonists_postop, na.rm = T)),
            SD = round(sd(Agonists_postop, na.rm = T)))
Agonist_postop

```

Calculate hours between assessment and last dopamine agonists taken: reported under section 10.1

```{r, message = FALSE, warning = FALSE}

last_agonist_preop <- data %>% 
  summarise(M = round(mean(last_agonist_preop, na.rm = T)),
            SD = round(sd(last_agonist_preop, na.rm = T)))
last_agonist_preop

last_agonist_postop <- data %>% 
  summarise(M = round(mean(last_agonist_postop, na.rm = T)),
            SD = round(sd(last_agonist_postop, na.rm = T)))
last_agonist_postop
```

Calculate percentage change of drug amount: reported under section 10.1

```{r, message = FALSE, warning = FALSE}

data$LEDD_change <- (data$LEDD_preop - data$LEDD_postop)/data$LEDD_preop * 100
data$agonist_change <- (data$Agonists_preop - data$Agonists_postop)/data$Agonists_preop * 100


LEDD_perc_change <- data %>% 
  summarise(M = round(mean(LEDD_change, na.rm = T)),
            SD = round(sd(LEDD_change, na.rm = T)))
LEDD_perc_change

Agonist_perc_change <- data %>% 
  summarise(M = round(mean(agonist_change, na.rm = T)),
            SD = round(sd(agonist_change, na.rm = T)))
Agonist_perc_change 

```