---
title: "Analysis of Vowel Space"
author: "Tabea Thies"
date: "`r Sys.Date()`"
output: html_document
---

# Analysis of Vowel Articulation Index (VAI) as measure of vowel space centralization

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load necessary packages

```{r load_tidyverse, warning = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(merTools)
library(sjPlot)

```

Load data set

```{r load_counts, message = FALSE, warning = FALSE}
data <- read_csv('../data/formants_VAI.csv')
```

# Create subset for the comparison Controls vs PD (in med-OFF)

```{r, message = FALSE, warning = FALSE}

data_CON_PD <-  filter(data, med_con %in% c('med-OFF', 'CON'))
```

Descriptive stats for reporting values in Section 6.2.2

```{r, message = FALSE, warning = FALSE}

VAI_stats <- data_CON_PD %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats

VAI_stats2 <- data_CON_PD %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats2

```

create linear mixed model 

```{r, message = FALSE, warning = FALSE}

set.seed(42)

mdl1 <- lmer(VAI_meas ~ med_con + focus_con + age + (1|speaker), data_CON_PD, REML = FALSE)

```

Look at model summary

```{r, message = FALSE, warning = FALSE}

summary(mdl1)
```

Extracting fit statistics for reporting in text

```{r, message = FALSE, warning = FALSE}
tab_model(mdl1)
```

## Explore random effects with function REsim

Effects for individual levels are simulated\
Output: mean, median and sd of the random effect estimates\

Plots estimated random effects for each participant\
95 % Intervals that do not overlap zero are bold

```{r, message = FALSE, warning = FALSE}
randoms <- REsim(mdl1) 
plotREsim(randoms)

```

Plot estimated 95 % interval spans for the intercept of VAI values

```{r, message = FALSE, warning = FALSE}

random2 <- subset(randoms, term == "(Intercept)")

random2 %>% 
  ggplot(aes(x = mean, xmin = mean-sd-sd, xmax = mean+sd+sd, y = reorder(groupID, mean))) +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  geom_point(size = 5)+
  geom_errorbar(size = 1) +
  theme_classic()+
  xlab('95 % intervals of estimated random intercepts for VAI') +
  ylab('Speaker')+
  theme(axis.title.y = element_text(face = 'bold', size = 18, margin = margin(r = 12)))+
  theme(axis.title.x = element_text(face = 'bold', size = 18, margin = margin(t = 12)))+
  theme(strip.text.x = element_text(size=18, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())+
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 18))
```


# Create subset for the comparison med-OFF vs med-ON

```{r, message = FALSE, warning = FALSE}

data_DOPA <-  filter(data, med_con %in% c('med-OFF', 'med-ON'))
```

Descriptive stats for reporting values in Section 6.2.2

```{r, message = FALSE, warning = FALSE}

VAI_stats <- data_DOPA %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats

VAI_stats2 <- data_DOPA %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats2

```

create linear mixed model 

```{r, message = FALSE, warning = FALSE}

set.seed(42)

mdl2 <- lmer(VAI_meas ~ med_con + focus_con + age + (1+med_con|speaker), data_DOPA, REML = FALSE)

```

Look at model summary

```{r, message = FALSE, warning = FALSE}

summary(mdl2)
```

Extracting fit statistics for reporting in text

```{r, message = FALSE, warning = FALSE}
tab_model(mdl2)
```

Explore random effects with function REsim

```{r, message = FALSE, warning = FALSE}
randoms <- REsim(mdl2) 
plotREsim(randoms)

```

Plot estimated 95 % interval spans for the intercept of VAI values

```{r, message = FALSE, warning = FALSE}

random2 <- subset(randoms, term == "(Intercept)")

random2 %>% 
  ggplot(aes(x = mean, xmin = mean-sd-sd, xmax = mean+sd+sd, y = reorder(groupID, mean))) +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  geom_point(size = 5)+
  geom_errorbar(size = 1) +
  theme_classic()+
  xlab('95 % intervals of estimated random intercepts for VAI') +
  ylab('Speaker')+
  theme(axis.title.y = element_text(face = 'bold', size = 18, margin = margin(r = 12)))+
  theme(axis.title.x = element_text(face = 'bold', size = 18, margin = margin(t = 12)))+
  theme(strip.text.x = element_text(size=18, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())+
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 18))
```

Plot estimated 95 % interval spans for the slopes of VAI values

```{r, message = FALSE, warning = FALSE}

random2 <- subset(randoms, term == "(Intercept)")

random2 %>% 
  ggplot(aes(x = mean, xmin = mean-sd-sd, xmax = mean+sd+sd, y = reorder(groupID, mean))) +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  geom_point(size = 5)+
  geom_errorbar(size = 1) +
  theme_classic()+
  xlab('95 % intervals of estimated random slopes for VAI') +
  ylab('Speaker')+
  theme(axis.title.y = element_text(face = 'bold', size = 18, margin = margin(r = 12)))+
  theme(axis.title.x = element_text(face = 'bold', size = 18, margin = margin(t = 12)))+
  theme(strip.text.x = element_text(size=18, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())+
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 18))
```

# Create subset for the comparison med-OFF vs med-ON

```{r, message = FALSE, warning = FALSE}

data_DOPA <-  filter(data, med_con %in% c('med-OFF', 'med-ON'))
```

Descriptive stats for reporting values in Section 7.2.2

```{r, message = FALSE, warning = FALSE}

VAI_stats <- data_DOPA %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats

VAI_stats2 <- data_DOPA %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats2

```

create linear mixed model 

```{r, message = FALSE, warning = FALSE}

set.seed(42)

mdl2 <- lmer(VAI_meas ~ med_con + focus_con + age + (1+med_con|speaker), data_DOPA, REML = FALSE)

```

Look at model summary

```{r, message = FALSE, warning = FALSE}

summary(mdl2)
```

Extracting fit statistics for reporting in text

```{r, message = FALSE, warning = FALSE}
tab_model(mdl2)
```

# Create subset for the comparison DBS-OFF vs DBS-ON

```{r, message = FALSE, warning = FALSE}

data_DBS <-  filter(data, med_con %in% c('med-OFF/DBS-OFF', 'med-OFF/DBS-ON'))

data_DBS <- subset(data_DBS, speaker != "PD03")

```

Descriptive stats for reporting values in Section 8.2.2

```{r, message = FALSE, warning = FALSE}

VAI_stats <- data_DBS %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats

VAI_stats2 <- data_DBS %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats2

```

create linear mixed model 

```{r, message = FALSE, warning = FALSE}

set.seed(42)

mdl3 <- lmer(VAI_meas ~ med_con + focus_con + age + (1+med_con|speaker), data_DBS, REML = FALSE)

```

Look at model summary

```{r, message = FALSE, warning = FALSE}

summary(mdl3)
```

Extracting fit statistics for reporting in text

```{r, message = FALSE, warning = FALSE}
tab_model(mdl3)
```

# Create subset for the comparison med-OFF vs DBS-OFF

```{r, message = FALSE, warning = FALSE}

data_OFF <-  filter(data, med_con %in% c('med-OFF', 'med-OFF/DBS-OFF'))

data_OFF <- subset(data_OFF, speaker != "PD03")

```

Descriptive stats for reporting values in Section 10.2.2

```{r, message = FALSE, warning = FALSE}

VAI_stats <- data_OFF %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats

VAI_stats2 <- data_OFF %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats2

```

create linear mixed model 

```{r, message = FALSE, warning = FALSE}

set.seed(42)

mdl4 <- lmer(VAI_meas ~ med_con + focus_con + age + (1+med_con|speaker), data_OFF, REML = FALSE)

```

Look at model summary

```{r, message = FALSE, warning = FALSE}

summary(mdl4)
```

Extracting fit statistics for reporting in text

```{r, message = FALSE, warning = FALSE}
tab_model(mdl4)
```

# Create subset for the comparison med-OFF vs DBS-ON
for the analysis if DOPA and DBS effects are different to baseline (med-OFF)

```{r, message = FALSE, warning = FALSE}

data_DBS2 <-  filter(data, med_con %in% c('med-OFF', 'med-OFF/DBS-ON'))

```

Descriptive stats for reporting values in Section 9.2.2

```{r, message = FALSE, warning = FALSE}

VAI_stats <- data_DBS2 %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats

VAI_stats2 <- data_DBS2 %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats2

```

create linear mixed model 

```{r, message = FALSE, warning = FALSE}

set.seed(42)

mdl5 <- lmer(VAI_meas ~ med_con + focus_con + age + (1+med_con|speaker), data_DBS2, REML = FALSE)

```

Look at model summary

```{r, message = FALSE, warning = FALSE}

summary(mdl5)
```

Extracting fit statistics for reporting in text

```{r, message = FALSE, warning = FALSE}
tab_model(mdl5)
```

