---
title: "Analysis of Vowel Space"
author: "Tabea Thies"
date: "`r Sys.Date()`"
output: html_document
---

# Analysis of Vowel Articulation Index (VAI) as measure of vowel space centralization

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load necessary packages

```{r load_tidyverse, warning = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(merTools)
library(sjPlot)
library(emmeans)

set.seed(42)

```

Load data set

```{r load_counts, message = FALSE, warning = FALSE}
data <- read_csv('../data/formants_VAI.csv')
```

# Create subset for the comparison Controls vs PD (in med-OFF)

```{r, message = FALSE, warning = FALSE}

data_CON_PD <-  filter(data, med_con %in% c('med-OFF', 'CON'))
```

Descriptive stats for reporting values in Section 6.2.2

```{r, message = FALSE, warning = FALSE}

VAI_stats <- data_CON_PD %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats

VAI_stats2 <- data_CON_PD %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats2

```

create linear mixed model 

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(VAI_meas ~ med_con * focus_con + (1|speaker), data_CON_PD, REML = FALSE)

mdl2 <- lmer(VAI_meas ~ med_con + focus_con + (1|speaker), data_CON_PD, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(VAI_meas ~ med_con + 1 + (1|speaker), data_CON_PD, REML = FALSE)

anova(mdl2, mdl2b) ## effect of focus: X2(2) = 10.758, p = 0.004613

emmeans(mdl2a, specs = pairwise ~ focus_con)

## check group

mdl2c <- lmer(VAI_meas ~ 1 + focus_con + (1|speaker), data_CON_PD, REML = FALSE)

anova(mdl2, mdl2c) ## no group effect: X2(1) = 1.4369, p = 0.2306

```

# Create subset for the comparison med-OFF vs med-ON

```{r, message = FALSE, warning = FALSE}

data_DOPA <-  filter(data, med_con %in% c('med-OFF', 'med-ON'))
```

Descriptive stats for reporting values in Section 6.2.2

```{r, message = FALSE, warning = FALSE}

VAI_stats <- data_DOPA %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats

VAI_stats2 <- data_DOPA %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats2

```

create linear mixed model 

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(VAI_meas ~ med_con * focus_con + (1|speaker), data_DOPA, REML = FALSE)

mdl2 <- lmer(VAI_meas ~ med_con + focus_con + (1|speaker), data_DOPA, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(VAI_meas ~ med_con + 1 + (1|speaker), data_DOPA, REML = FALSE)

anova(mdl2, mdl2b) ## no effect of focus: X2(2) = 3.9446, p = 0.1391

## check group

mdl2c <- lmer(VAI_meas ~ 1 + focus_con + (1|speaker), data_DOPA, REML = FALSE)

anova(mdl2, mdl2c) ## no group effect: X2(1) = 0.2405, p = 0.6238


```

# Create subset for the comparison DBS-OFF vs DBS-ON

```{r, message = FALSE, warning = FALSE}

data_DBS <-  filter(data, med_con %in% c('med-OFF/DBS-OFF', 'med-OFF/DBS-ON'))

data_DBS <- subset(data_DBS, speaker != "PD03")

```

Descriptive stats for reporting values in Section 8.2.2

```{r, message = FALSE, warning = FALSE}

VAI_stats <- data_DBS %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats

VAI_stats2 <- data_DBS %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats2

```

create linear mixed model 

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(VAI_meas ~ med_con * focus_con + (1|speaker), data_DBS, REML = FALSE)

mdl2 <- lmer(VAI_meas ~ med_con + focus_con + (1|speaker), data_DBS, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(VAI_meas ~ med_con + 1 + (1|speaker), data_DBS, REML = FALSE)

anova(mdl2, mdl2b) ## no effect of focus: X2(2) = 1.9426 , p = 0.3786

## check group

mdl2c <- lmer(VAI_meas ~ 1 + focus_con + (1|speaker), data_DBS, REML = FALSE)

anova(mdl2, mdl2c) ## no group effect: X2(1) = 0.4425, p = 0.5059

```

# Create subset for the comparison med-OFF vs DBS-OFF

```{r, message = FALSE, warning = FALSE}

data_OFF <-  filter(data, med_con %in% c('med-OFF', 'med-OFF/DBS-OFF'))

data_OFF <- subset(data_OFF, speaker != "PD03")

```

Descriptive stats for reporting values in Section 10.2.2

```{r, message = FALSE, warning = FALSE}

VAI_stats <- data_OFF %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats

VAI_stats2 <- data_OFF %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats2

```

create linear mixed model 

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(VAI_meas ~ med_con * focus_con + (1|speaker), data_OFF, REML = FALSE)

mdl2 <- lmer(VAI_meas ~ med_con + focus_con + (1|speaker), data_OFF, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(VAI_meas ~ med_con + 1 + (1|speaker), data_OFF, REML = FALSE)

anova(mdl2, mdl2b) ## no effect of focus: X2(2) = 3.1722, p = 0.2047


## check group

mdl2c <- lmer(VAI_meas ~ 1 + focus_con + (1|speaker), data_OFF, REML = FALSE)

anova(mdl2, mdl2c) ## no group effect: X2(1) = 0.3209 , p = 0.5711

```

# Create subset for the comparison med-OFF vs DBS-ON
for the analysis if DOPA and DBS effects are different to baseline (med-OFF)

```{r, message = FALSE, warning = FALSE}

data_DBS2 <-  filter(data, med_con %in% c('med-OFF', 'med-ON', 'med-OFF/DBS-ON'))

```

Descriptive stats for reporting values in Section 9.2.2

```{r, message = FALSE, warning = FALSE}

VAI_stats <- data_DBS2 %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats

VAI_stats2 <- data_DBS2 %>% 
  group_by(med_con, focus_con) %>% 
  summarise(M = round(mean(VAI_meas,na.rm=T),2),
            SD = round(sd(VAI_meas,na.rm=T), 2))
VAI_stats2

```

create linear mixed model 

```{r, message = FALSE, warning = FALSE}

## interaction 
mdl1 <- lmer(VAI_meas ~ med_con * focus_con + (1|speaker), data_DBS2, REML = FALSE)

mdl2 <- lmer(VAI_meas ~ med_con + focus_con + (1|speaker), data_DBS2, REML = FALSE)

anova(mdl1, mdl2) ## no interaction

## check focus

mdl2b <- lmer(VAI_meas ~ med_con + 1 + (1|speaker), data_DBS2, REML = FALSE)

anova(mdl2, mdl2b) ## no effect of focus: X2(2) = 4.1614, p = 0.1248


## check group

mdl2c <- lmer(VAI_meas ~ 1 + focus_con + (1|speaker), data_DBS2, REML = FALSE)

anova(mdl2, mdl2c) ## no group effect: X2(1) = 0.2806, p = 0.8691

```