---
title: "Analysis of VAS Ratings"
author: "Tabea Thies"
date: "`r Sys.Date()`"
output: html_document
---

# Analysis of naive listener ratings of intelligibility and naturalness

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load necessary packages

```{r load_tidyverse, warning = FALSE, message = FALSE}

library(tidyverse)
library(ordinalCont)

set.seed(42)

```

Load data set

```{r load_counts, message = FALSE, warning = FALSE}
data <- read_csv('../data/listener_ratings.csv')
```

# Create subsets for relevant comparisons

and exclude data of speaker PD03 because DBS-OFF condition is missing

```{r, message = FALSE, warning = FALSE}

data_PD <-  filter(data, Group %in% c('medOFF', 'CON'))

data_DOPA <-  filter(data, Group %in% c('medOFF', 'medON'))

data_DBS <-  filter(data, Group %in% c('DBSOFF', 'DBSON'))
data_DBS <- subset(data_DBS, Speaker != "PD03")

data_OFF <-  filter(data, Group %in% c('medOFF', 'DBSOFF'))
data_OFF <- subset(data_OFF, Speaker != "PD03")

data_DBS2 <-  filter(data, Group %in% c('medOFF', 'DBSON'))

```

# Disease Effect

Descriptive stats for reporting values in Section 6.2.1

```{r, message = FALSE, warning = FALSE}

Intell <- data_PD %>% 
  group_by(Group) %>% 
  summarise(M = round(mean(Intell)),
            SD = round(sd(Intell)))

Intell

Natural <- data_PD %>% 
  group_by(Group) %>% 
  summarise(M = round(mean(Natural)),
            SD = round(sd(Natural)))
Natural

```

create continuous ordinal regression model
for testing difference of intelligibility 

```{r, message = FALSE, warning = FALSE}

#fit1 <- ocm(Intell ~ Group + vowel + cons + (1|Speaker) + (1|Rater),
#            data = data_PD)

#save(fit1, file = '../models/Intell_mdl_PD.RData')

load('../models/Intell_mdl_PD.RData')

summary(fit1)
```

create continuous ordinal regression model
for testing difference of naturalness 

```{r, message = FALSE, warning = FALSE}

#fit2 <- ocm(Natural ~ Group + vowel + cons + (1|Speaker) + (1|Rater),
#            data = data_PD)

#save(fit2, file = '../models/Natural_mdl_PD.RData')

load('../models/Natural_mdl_PD.RData')

summary(fit2)

```

# Levodopa Effect

Descriptive stats for reporting values in Section 7.2.1

```{r, message = FALSE, warning = FALSE}

Intell <- data_DOPA %>% 
  group_by(Group) %>% 
  summarise(M = round(mean(Intell)),
            SD = round(sd(Intell)))

Intell

Natural <- data_DOPA %>% 
  group_by(Group) %>% 
  summarise(M = round(mean(Natural)),
            SD = round(sd(Natural)))
Natural

```

create continuous ordinal regression model
for testing difference of intelligibility 

```{r, message = FALSE, warning = FALSE}

#fit3 <- ocm(Intell ~ Group + vowel + cons + (1|Speaker) + (1|Rater),
#            data = data_DOPA)

#save(fit3, file = '../models/Intell_mdl_DOPA.RData')

load('../models/Intell_mdl_DOPA.RData')

summary(fit3)

```

create continuous ordinal regression model
for testing difference of naturalness 

```{r, message = FALSE, warning = FALSE}

#fit4 <- ocm(Natural ~ Group + vowel + cons + (1|Speaker) + (1|Rater),
#            data = data_DOPA)

#save(fit4, file = '../models/Natural_mdl_DOPA.RData')

load('../models/Natural_mdl_DOPA.RData')

summary(fit4)

```

# DBS Effect

Descriptive stats for reporting values in Section 8.2.1

```{r, message = FALSE, warning = FALSE}

Intell <- data_DBS %>% 
  group_by(Group) %>% 
  summarise(M = round(mean(Intell)),
            SD = round(sd(Intell)))

Intell

Natural <- data_DBS %>% 
  group_by(Group) %>% 
  summarise(M = round(mean(Natural)),
            SD = round(sd(Natural)))
Natural

```

create continuous ordinal regression model
for testing difference of intelligibility 

```{r, message = FALSE, warning = FALSE}

#fit5 <- ocm(Intell ~ Group + vowel + cons + (1|Speaker) + (1|Rater),
#            data = data_DBS)

#save(fit5, file = '../models/Intell_mdl_DBS.RData')

load('../models/Intell_mdl_DBS.RData')

summary(fit5)

```

create continuous ordinal regression model
for testing difference of naturalness 

```{r, message = FALSE, warning = FALSE}

#fit6 <- ocm(Natural ~ Group + vowel + cons + (1|Speaker) + (1|Rater),
#            data = data_DBS)

#save(fit6, file = '../models/Natural_mdl_DBS.RData')

load('../models/Natural_mdl_DBS.RData')

summary(fit6)

```

# Electrode Effect

Descriptive stats for reporting values in Section 10.2.1

```{r, message = FALSE, warning = FALSE}

Intell <- data_OFF %>% 
  group_by(Group) %>% 
  summarise(M = round(mean(Intell)),
            SD = round(sd(Intell)))

Intell

Natural <- data_OFF %>% 
  group_by(Group) %>% 
  summarise(M = round(mean(Natural)),
            SD = round(sd(Natural)))
Natural

```

create continuous ordinal regression model
for testing difference of intelligibility 

```{r, message = FALSE, warning = FALSE}

#fit7 <- ocm(Intell ~ Group + vowel + cons + (1|Speaker) + (1|Rater),
#            data = data_OFF)

#save(fit7, file = '../models/Intell_mdl_OFF.RData')

load('../models/Intell_mdl_OFF.RData')

summary(fit7)

```

create continuous ordinal regression model
for testing difference of naturalness 

```{r, message = FALSE, warning = FALSE}

#fit8 <- ocm(Natural ~ Group + vowel + cons + (1|Speaker) + (1|Rater),
#            data = data_OFF)

#save(fit8, file = '../models/Natural_mdl_OFF.RData')

load('../models/Natural_mdl_OFF.RData')

summary(fit8)


```

# Comparision: Baseline (preop med-OFF) with DBS-ON

Descriptive stats for reporting values in Section 9.2.1

```{r, message = FALSE, warning = FALSE}

Intell <- data_DBS2 %>% 
  group_by(Group) %>% 
  summarise(M = round(mean(Intell)),
            SD = round(sd(Intell)))

Intell

Natural <- data_DBS2 %>% 
  group_by(Group) %>% 
  summarise(M = round(mean(Natural)),
            SD = round(sd(Natural)))
Natural

```

create continuous ordinal regression model
for testing difference of intelligibility 

```{r, message = FALSE, warning = FALSE}

#fit9 <- ocm(Intell ~ Group + vowel + cons + (1|Speaker) + (1|Rater),
#            data = data_DBS2)

#save(fit9, file = '../models/Intell_mdl_DBS2.RData')

load('../models/Intell_mdl_DBS2.RData')

summary(fit9)

```

create continuous ordinal regression model
for testing difference of naturalness 

```{r, message = FALSE, warning = FALSE}

#fit10 <- ocm(Natural ~ Group + vowel + cons + (1|Speaker) + (1|Rater),
#            data = data_DBS2)

#save(fit10, file = '../models/Natural_mdl_DBS2.RData')

load('../models/Natural_mdl_DBS2.RData')

summary(fit10)

```
