---
title: "General Information"
author: "Tabea Thies"
date: "`r Sys.Date()`"
output: html_document
---

# Analysis of motor and cognitive scores

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load necessary packages

```{r load_tidyverse, warning = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(lme4)
set.seed(42)
library(ordinalCont)
```

Load data set

```{r load_counts, message = FALSE, warning = FALSE}
data <- read_delim('../data/demographics.csv')
```

## Disease duration

```{r, message = FALSE, warning = FALSE}

data_dis_dur <-  filter(data, med_con %in% c('med-OFF'))

disdur_avgs <- data_dis_dur %>% 
  summarise(M = round(mean(dis_dur, na.rm = T)),
            SD = round(sd(dis_dur, na.rm = T)))
disdur_avgs

```
## Age

```{r, message = FALSE, warning = FALSE}

data_age <-  filter(data, med_con %in% c('CON', 'med-OFF'))

age_avgs <- data_age %>% 
    group_by(med_con) %>% 
    summarise(M = round(mean(age, na.rm = T)),
            SD = round(sd(age, na.rm = T)))
age_avgs

```

# UPDRS analysis

Disease effect: comparing motor scores between controls and PD in med-OFF
reported under section 6.1

```{r, message = FALSE, warning = FALSE}

data_CON_PD <-  filter(data, med_con %in% c('CON', 'med-OFF'))

UPDRS_avgs <- data_CON_PD %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(UPDRS_total, na.rm = T)),
            SD = round(sd(UPDRS_total, na.rm = T)),
            min = min(UPDRS_total, na.rm = T),
            max = max(UPDRS_total, na.rm = T))
UPDRS_avgs

mdl1 <- lm(UPDRS_total ~ med_con  + age, data_CON_PD, REML = FALSE)
mdl1_null <- lm(UPDRS_total ~ 1 + age, data_CON_PD, REML = FALSE)

anova(mdl1, mdl1_null)

axial_stats <- data_CON_PD %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(UPDRS_axial, na.rm = T)),
            SD = round(sd(UPDRS_axial, na.rm = T)))
axial_stats

brady_stats <- data_CON_PD %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(UPDRS_brady, na.rm = T)),
            SD = round(sd(UPDRS_brady, na.rm = T)))
brady_stats

rigid_stats <- data_CON_PD %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(UPDRS_rigidity, na.rm = T)),
            SD = round(sd(UPDRS_rigidity, na.rm = T)))
rigid_stats 

PIGD_stats <- data_CON_PD %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(UPDRS_PIGD, na.rm = T)),
            SD = round(sd(UPDRS_PIGD, na.rm = T)))
PIGD_stats

```

Levodopa effect: comparing motor scores between med-OFF and med-ON
reported and plotted under section 7.1

```{r, message = FALSE, warning = FALSE}

data_preop <-  filter(data, med_con %in% c('med-OFF', 'med-ON'))

UPDRS_avgs <- data_preop %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(UPDRS_total, na.rm = T)),
            SD = round(sd(UPDRS_total, na.rm = T)),
            min = min(UPDRS_total, na.rm = T),
            max = max(UPDRS_total, na.rm = T))
UPDRS_avgs

mdl2 <- lm(UPDRS_total ~ med_con  + age, data_preop, REML = FALSE)
mdl2_null <- lm(UPDRS_total ~ 1 + age, data_preop, REML = FALSE)

anova(mdl2, mdl2_null)

UPDRS_avgs2 <- data_preop %>% 
  group_by(med_con, speaker) %>% 
  summarise(M = round(mean(UPDRS_total, na.rm = T)),
            SD = round(sd(UPDRS_total, na.rm = T)))

UPDRS_avgs2 %>% 
  ggplot(aes(x = med_con, y = M, group = speaker)) +
  geom_point(aes(col = med_con), shape = 19, alpha = 0.7, size = 6) +
  geom_line(col = 'darkgrey') +
  theme_classic() +
  theme(legend.position = 'none')+
  scale_color_manual(values = c('#BE7592', '#74BDA0')) +
  xlab(NULL) +
  scale_x_discrete(labels = c('med-OFF','med-ON'))+
  ylab("UPDRS III")+
  theme(axis.title.y = element_text(face = 'bold', size = 18, margin = margin(r = 12)))+
  theme(strip.text.x = element_text(size=20, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())
```

DBS effect: comparing motor scores between DBS-OFF and DBS-ON
reported and plotted under section 8.1

```{r, message = FALSE, warning = FALSE}

data_postop <-  filter(data, med_con %in% c('med-OFF/DBS-OFF', 'med-OFF/DBS-ON'))

## remove data of PD03
data_postop <- data_postop[-c(5,6), ]

UPDRS_avgs <- data_postop %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(UPDRS_total, na.rm = T)),
            SD = round(sd(UPDRS_total, na.rm = T)),
            min = min(UPDRS_total, na.rm = T),
            max = max(UPDRS_total, na.rm = T))
UPDRS_avgs

mdl3 <- lm(UPDRS_total ~ med_con  + age, data_postop, REML = FALSE)
mdl3_null <- lm(UPDRS_total ~ 1 + age, data_postop, REML = FALSE)

anova(mdl3, mdl3_null)

UPDRS_avgs2 <- data_postop %>% 
  group_by(med_con, speaker) %>% 
  summarise(M = round(mean(UPDRS_total, na.rm = T)),
            SD = round(sd(UPDRS_total, na.rm = T)))

UPDRS_avgs2 %>% 
  ggplot(aes(x = med_con, y = M, group = speaker)) +
  geom_point(aes(col = med_con), shape = 19, alpha = 0.7, size = 6) +
  geom_line(col = 'darkgrey') +
  theme_classic() +
  theme(legend.position = 'none')+
  scale_color_manual(values = c('#5275AD', '#AD8A52')) +
  xlab(NULL) +
  scale_x_discrete(labels = c('DBS-OFF','DBS-ON'))+
  ylab("UPDRS III")+
  theme(axis.title.y = element_text(face = 'bold', size = 18, margin = margin(r = 12)))+
  theme(strip.text.x = element_text(size=20, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())
```

Electrode/OFF effect: comparing motor scores between med-OFF and DBS-OFF
reported and plotted under section 10.1

```{r, message = FALSE, warning = FALSE}

data_OFF <-  filter(data, med_con %in% c('med-OFF', 'med-OFF/DBS-OFF'))

## remove data of PD03
data_OFF <- data_OFF[-c(5,6), ]

UPDRS_avgs <- data_OFF %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(UPDRS_total, na.rm = T)),
            SD = round(sd(UPDRS_total, na.rm = T)),
            min = min(UPDRS_total, na.rm = T),
            max = max(UPDRS_total, na.rm = T))
UPDRS_avgs

mdl4 <- lm(UPDRS_total ~ med_con  + age, data_OFF, REML = FALSE)
mdl4_null <- lm(UPDRS_total ~ 1 + age, data_OFF, REML = FALSE)

anova(mdl4, mdl4_null)

UPDRS_avgs2 <- data_OFF %>% 
  group_by(med_con, speaker) %>% 
  summarise(M = round(mean(UPDRS_total, na.rm = T)),
            SD = round(sd(UPDRS_total, na.rm = T)))

UPDRS_avgs2 %>% 
  ggplot(aes(x = med_con, y = M, group = speaker)) +
  geom_point(aes(col = med_con), shape = 19, alpha = 0.7, size = 6) +
  geom_line(col = 'darkgrey') +
  theme_classic() +
  theme(legend.position = 'none')+
  scale_color_manual(values = c('#BE7592', '#5275AD')) +
  xlab(NULL) +
  scale_x_discrete(labels = c('med-OFF','DBS-OFF'))+
  ylab("UPDRS III")+
  theme(axis.title.y = element_text(face = 'bold', size = 18, margin = margin(r = 12)))+
  theme(strip.text.x = element_text(size=20, face="bold"),
        strip.background = element_blank())+
  theme(axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18)) +
  theme(axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank())
```

# Percentage changes in motor scores
Create subsets

```{r, message = FALSE, warning = FALSE}

data_medOFF <-  filter(data, med_con %in% c('med-OFF'))
data_medON <-  filter(data, med_con %in% c('med-ON'))
data_DBSOFF <-  filter(data, med_con %in% c('med-OFF/DBS-OFF'))
data_DBSON <-  filter(data, med_con %in% c('med-OFF/DBS-ON'))

```

Calculate change in motoric subscores from med-OFF to med-ON
reported under 7.1

```{r, message = FALSE, warning = FALSE}

UPDRS_OFF_avgs <- data_medOFF %>% 
  group_by(speaker) %>% 
  summarise(UPDRS_medOFF = mean(UPDRS_total, na.rm = T),
            axial_medOFF = mean(UPDRS_axial, na.rm = T),
            rigid_medOFF = mean(UPDRS_rigidity, na.rm = T),
            brady_medOFF = mean(UPDRS_brady, na.rm = T),
            PIGD_medOFF = mean(UPDRS_PIGD, na.rm = T))

UPDRS_ON_avgs <- data_medON  %>% 
  group_by(speaker) %>% 
  summarise(UPDRS_medON = mean(UPDRS_total, na.rm = T),
            axial_medON = mean(UPDRS_axial, na.rm = T),
            rigid_medON = mean(UPDRS_rigidity, na.rm = T),
            brady_medON = mean(UPDRS_brady, na.rm = T),
            PIGD_medON = mean(UPDRS_PIGD, na.rm = T))

UPDRS_change <- inner_join(UPDRS_OFF_avgs, UPDRS_ON_avgs, by="speaker")

## calculate percentage change

UPDRS_change$total_change <- (UPDRS_change$UPDRS_medOFF - UPDRS_change$UPDRS_medON)/UPDRS_change$UPDRS_medOFF * 100
UPDRS_change$axial_change <- (UPDRS_change$axial_medOFF - UPDRS_change$axial_medON)/UPDRS_change$axial_medOFF * 100
UPDRS_change$rigid_change <- (UPDRS_change$rigid_medOFF - UPDRS_change$rigid_medON)/UPDRS_change$rigid_medOFF * 100
UPDRS_change$PIGD_change <- (UPDRS_change$PIGD_medOFF - UPDRS_change$PIGD_medON)/UPDRS_change$PIGD_medOFF * 100
UPDRS_change$brady_change <- (UPDRS_change$brady_medOFF - UPDRS_change$brady_medON)/UPDRS_change$brady_medOFF * 100


UPDRS_perc_change <- UPDRS_change %>% 
  summarise(M = round(mean(total_change, na.rm = T)),
            SD = round(sd(total_change, na.rm = T)))
UPDRS_perc_change

axial_perc_change <- UPDRS_change %>% 
  summarise(M = round(mean(axial_change, na.rm = T)),
            SD = round(sd(axial_change, na.rm = T)))
axial_perc_change

rigid_perc_change <- UPDRS_change %>% 
  summarise(M = round(mean(rigid_change, na.rm = T)),
            SD = round(sd(rigid_change, na.rm = T)))
rigid_perc_change

UPDRS_change$PIGD_change[UPDRS_change$PIGD_change == -Inf] <- NA

PIGD_perc_change <- UPDRS_change %>% 
  summarise(M = round(mean(PIGD_change, na.rm = T)),
            SD = round(sd(PIGD_change, na.rm = T)))
PIGD_perc_change 

UPDRS_change$brady_change[UPDRS_change$brady_change == -Inf] <- NA

brady_perc_change <- UPDRS_change %>% 
  summarise(M = round(mean(brady_change, na.rm = T)),
            SD = round(sd(brady_change, na.rm = T)))
brady_perc_change 

```

Calculate change in motoric subscores from DBS-OFF to DBS-ON
reported under 8.1

```{r, message = FALSE, warning = FALSE}

UPDRS_OFF_avgs <- data_DBSOFF %>% 
  group_by(speaker) %>% 
  summarise(UPDRS_DBSOFF = mean(UPDRS_total, na.rm = T),
            axial_DBSOFF = mean(UPDRS_axial, na.rm = T),
            rigid_DBSOFF = mean(UPDRS_rigidity, na.rm = T),
            brady_DBSOFF = mean(UPDRS_brady, na.rm = T),
            PIGD_DBSOFF = mean(UPDRS_PIGD, na.rm = T))

UPDRS_ON_avgs <- data_DBSON  %>% 
  group_by(speaker) %>% 
  summarise(UPDRS_DBSON = mean(UPDRS_total, na.rm = T),
            axial_DBSON = mean(UPDRS_axial, na.rm = T),
            rigid_DBSON = mean(UPDRS_rigidity, na.rm = T),
            brady_DBSON = mean(UPDRS_brady, na.rm = T),
            PIGD_DBSON = mean(UPDRS_PIGD, na.rm = T))

UPDRS_change <- inner_join(UPDRS_OFF_avgs, UPDRS_ON_avgs, by="speaker")

## remove data of PD03
UPDRS_change <- UPDRS_change[-3, ]

## calculate percentage change

UPDRS_change$total_change <- (UPDRS_change$UPDRS_DBSOFF - UPDRS_change$UPDRS_DBSON)/UPDRS_change$UPDRS_DBSOFF * 100
UPDRS_change$axial_change <- (UPDRS_change$axial_DBSOFF - UPDRS_change$axial_DBSON)/UPDRS_change$axial_DBSOFF * 100
UPDRS_change$rigid_change <- (UPDRS_change$rigid_DBSOFF - UPDRS_change$rigid_DBSON)/UPDRS_change$rigid_DBSOFF * 100
UPDRS_change$PIGD_change <- (UPDRS_change$PIGD_DBSOFF - UPDRS_change$PIGD_DBSON)/UPDRS_change$PIGD_DBSOFF * 100
UPDRS_change$brady_change <- (UPDRS_change$brady_DBSOFF - UPDRS_change$brady_DBSON)/UPDRS_change$brady_DBSOFF * 100

UPDRS_perc_change <- UPDRS_change %>% 
  summarise(M = round(mean(total_change, na.rm = T)),
            SD = round(sd(total_change, na.rm = T)))
UPDRS_perc_change

axial_perc_change <- UPDRS_change %>% 
  summarise(M = round(mean(axial_change, na.rm = T)),
            SD = round(sd(axial_change, na.rm = T)))
axial_perc_change

rigid_perc_change <- UPDRS_change %>% 
  summarise(M = round(mean(rigid_change, na.rm = T)),
            SD = round(sd(rigid_change, na.rm = T)))
rigid_perc_change

PIGD_perc_change <- UPDRS_change %>% 
  summarise(M = round(mean(PIGD_change, na.rm = T)),
            SD = round(sd(PIGD_change, na.rm = T)))
PIGD_perc_change 

brady_perc_change <- UPDRS_change %>% 
  summarise(M = round(mean(brady_change, na.rm = T)),
            SD = round(sd(brady_change, na.rm = T)))
brady_perc_change 

```


Calculate change in motoric subscores from med-OFF to DBS-ON
reported under 9.1

```{r, message = FALSE, warning = FALSE}

UPDRS_OFF_avgs <- data_medOFF %>% 
  group_by(speaker) %>% 
  summarise(UPDRS_medOFF = mean(UPDRS_total, na.rm = T),
            axial_medOFF = mean(UPDRS_axial, na.rm = T),
            rigid_medOFF = mean(UPDRS_rigidity, na.rm = T),
            brady_medOFF = mean(UPDRS_brady, na.rm = T),
            PIGD_medOFF = mean(UPDRS_PIGD, na.rm = T))

UPDRS_ON_avgs <- data_DBSON  %>% 
  group_by(speaker) %>% 
  summarise(UPDRS_DBSON = mean(UPDRS_total, na.rm = T),
            axial_DBSON = mean(UPDRS_axial, na.rm = T),
            rigid_DBSON = mean(UPDRS_rigidity, na.rm = T),
            brady_DBSON = mean(UPDRS_brady, na.rm = T),
            PIGD_DBSON = mean(UPDRS_PIGD, na.rm = T))

UPDRS_change <- inner_join(UPDRS_OFF_avgs, UPDRS_ON_avgs, by="speaker")

## calculate percentage change

UPDRS_change$total_change <- (UPDRS_change$UPDRS_medOFF - UPDRS_change$UPDRS_DBSON)/UPDRS_change$UPDRS_medOFF * 100
UPDRS_change$axial_change <- (UPDRS_change$axial_medOFF - UPDRS_change$axial_DBSON)/UPDRS_change$axial_medOFF * 100
UPDRS_change$rigid_change <- (UPDRS_change$rigid_medOFF - UPDRS_change$rigid_DBSON)/UPDRS_change$rigid_medOFF * 100
UPDRS_change$PIGD_change <- (UPDRS_change$PIGD_medOFF - UPDRS_change$PIGD_DBSON)/UPDRS_change$PIGD_medOFF * 100
UPDRS_change$brady_change <- (UPDRS_change$brady_medOFF - UPDRS_change$brady_DBSON)/UPDRS_change$brady_medOFF * 100

UPDRS_perc_change <- UPDRS_change %>% 
  summarise(M = round(mean(total_change, na.rm = T)),
            SD = round(sd(total_change, na.rm = T)))
UPDRS_perc_change

axial_perc_change <- UPDRS_change %>% 
  summarise(M = round(mean(axial_change, na.rm = T)),
            SD = round(sd(axial_change, na.rm = T)))
axial_perc_change

rigid_perc_change <- UPDRS_change %>% 
  summarise(M = round(mean(rigid_change, na.rm = T)),
            SD = round(sd(rigid_change, na.rm = T)))
rigid_perc_change

UPDRS_change$PIGD_change[UPDRS_change$PIGD_change == -Inf] <- NA

PIGD_perc_change <- UPDRS_change %>% 
  summarise(M = round(mean(PIGD_change, na.rm = T)),
            SD = round(sd(PIGD_change, na.rm = T)))
PIGD_perc_change 

UPDRS_change$brady_change[UPDRS_change$brady_change == -Inf] <- NA

brady_perc_change <- UPDRS_change %>% 
  summarise(M = round(mean(brady_change, na.rm = T)),
            SD = round(sd(brady_change, na.rm = T)))
brady_perc_change 

```

# VAS analysis

Disease effect: comparing controls and PD in med-OFF
reported under section 6.2.1

```{r, message = FALSE, warning = FALSE}

VAS_avgs <- data_CON_PD %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(VAS, na.rm = T),1),
            SD = round(sd(VAS, na.rm = T),1))
VAS_avgs

fit1 <- ocm(VAS ~ med_con + (1|speaker),
            data = data_CON_PD)

summary(fit1)

```

Levodopa effect: comparing med-OFF and med-ON
reported under section 7.2.1

```{r, message = FALSE, warning = FALSE}

VAS_avgs <- data_preop  %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(VAS, na.rm = T),1),
            SD = round(sd(VAS, na.rm = T),1))
VAS_avgs

fit2 <- ocm(VAS ~ med_con + (1|speaker),
            data = data_preop )

summary(fit2)

```

DBS effect: comparing DBS-OFF and DBS-ON
reported under section 8.2.1

```{r, message = FALSE, warning = FALSE}

VAS_avgs <- data_postop  %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(VAS, na.rm = T),1),
            SD = round(sd(VAS, na.rm = T),1))
VAS_avgs

fit3 <- ocm(VAS ~ med_con + (1|speaker),
            data = data_postop )

summary(fit3)

```


Electrode effect: comparing med-OFF and DBS-OFF
reported under section 10.2.1

```{r, message = FALSE, warning = FALSE}

VAS_avgs <- data_OFF  %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(VAS, na.rm = T),1),
            SD = round(sd(VAS, na.rm = T),1))
VAS_avgs

fit4 <- ocm(VAS ~ med_con + (1|speaker),
            data = data_OFF)

summary(fit4)

```

DBS effect to Baseline: comparing med-OFF and DBS-ON
reported under section .2.1

```{r, message = FALSE, warning = FALSE}

data_ON_DBS <-  filter(data, med_con %in% c('med-OFF', 'med-OFF/DBS-ON'))

VAS_avgs <- data_ON_DBS  %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(VAS, na.rm = T),1),
            SD = round(sd(VAS, na.rm = T),1))
VAS_avgs

fit5 <- ocm(VAS ~ med_con + (1|speaker),
            data = data_ON_DBS)

summary(fit5)

```

# Cognitive scores

averaged across group (PD vs CON) and visit (preop vs postop)

```{r, message = FALSE, warning = FALSE}

data_cog <-  filter(data, med_con %in% c('CON', 'med-ON', 'med-OFF/DBS-ON'))

BDI_stats <- data_cog %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(BDI, na.rm = T)),
            SD = round(sd(BDI, na.rm = T)))
BDI_stats

PANDA_stats <- data_cog %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(PANDA, na.rm = T)),
            SD = round(sd(PANDA, na.rm = T)))
PANDA_stats

MMST_stats <- data_cog %>% 
  group_by(med_con) %>% 
  summarise(M = round(mean(MMST, na.rm = T)),
            SD = round(sd(MMST, na.rm = T)))
MMST_stats

```

